---
title: "02_Dimred_excess"
author: "Kolja Nenoff"
date: "2023-06-11"
output:
  rmdformats::readthedown:
    code_folding: hide
    self_contained: true
    css: custom.css 
    lightbox: true
    gallery: true
    highlight: tango  
    toc_depth: 5
---


Test the linear relationship of dimred and the 


# 0 Preprocessing

## Libraries

```{r,include=T,message=FALSE}
rm(list = ls())
require(DT)
library(sf)
#remotes::install_github("coolbutuseless/ggpattern")
require(ggpattern)
require(ggpubr)
library(rnaturalearth)
library(rnaturalearthdata)
require(ggthemes)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(wesanderson)
library(tidyverse)
require(zoo)
library(sf)
require(data.table)
library(rnaturalearth)
library(rnaturalearthdata)
require(openxlsx)
require(readxl)
require(energy)
library(quantreg)
require(Boruta)
require(kableExtra)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
require(dplyr)
require(iml)

```
## Data

### 1. Excess Moratility
```{r,Excess Mortility WHO Set}
excessCounrty_year<-read_excel("../data/230224_nature_WHO_excess_death/2022-03-25_covid-19_gem/WHO_COVID_Excess_Deaths_EstimatesByCountry.xlsx",sheet = 2)
excessCounrty_covariates<-read_excel("../data/230224_nature_WHO_excess_death/2022-03-25_covid-19_gem/WHO_COVID_Excess_Deaths_EstimatesByCountry.xlsx",sheet = 3)
excessRegion_year<-read_excel("../data/230224_nature_WHO_excess_death/2022-03-25_covid-19_gem/WHO_COVID_Excess_Deaths_EstimatesByRegion.xlsx",sheet = 2)
excessRegion_income<-read_excel("../data/230224_nature_WHO_excess_death/2022-03-25_covid-19_gem/WHO_COVID_Excess_Deaths_EstimatesByIncome.xlsx",sheet = 2)
excessRegion_income %>% group_by(income) %>% 
  summarise(p_value_mort=sum(excess.mean)/sum(expected.mean)*100,excess_comul=sum(excess.mean))
excessCounrty_year_pvalue<-excessCounrty_year %>% group_by(country,iso3,year) %>% 
  summarise(p_value_mort=sum(excess.mean)/sum(expected.mean)*100,excess_comul=sum(excess.mean))
excessCounrty_year %>% filter(year==2020)-> excessCounrty_year_2020
p_value_global_2020=sum(excessCounrty_year_2020$excess.mean,na.rm = T)/sum(excessCounrty_year_2020$expected.mean,na.rm = T)*100
excessCounrty_year %>% filter(year==2021)-> excessCounrty_year_2020
p_value_global_2021=sum(excessCounrty_year_2020$excess.mean,na.rm = T)/sum(excessCounrty_year_2020$expected.mean,na.rm = T)*100
```


```{r,message=FALSE}
incomegroups<-read_excel("../data/230224_WORLDBANK_INCOME_GROUPS/CLASS.xlsx")
incomegroups2<-read_excel("../data/230224_WORLDBANK_INCOME_GROUPS/OGHIST.xlsx",sheet = 2) 
availability<-read.xlsx("../data/230320_WHO_Mortality_availability/41586_2022_5522_availability.xlsx")
wdi_explained<-fread("../data/WDI_data_230609/WDI_CSV/WDISeries.csv")


```



```{r,message=FALSE}

incomegroups$incomegroup<-incomegroups$`Income group`

excessCounrty_year_pvalue2<-merge(excessCounrty_year_pvalue,incomegroups,all.x = T,by.x="iso3",by.y="Code")

excessCounrty_year_pvalue2<-merge(excessCounrty_year_pvalue2,availability,by.x="iso3",by.y="iso3")

excessCounrty_year_pvalue2$identifier<-paste0(excessCounrty_year_pvalue2$iso3,"_",excessCounrty_year_pvalue2$year)
excessCounrty_year_pvalue2$iso3_code<-excessCounrty_year_pvalue2$iso3
excessCounrty_year_pvalue2$iso3<-NULL
excessCounrty_year_pvalue2$excess_comul<-NULL
excessCounrty_year_pvalue2$identifier<-NULL
excessCounrty_year_pvalue2 %>% filter(!is.na(iso3_code))%>% pivot_wider(names_from = year, values_from = p_value_mort) %>% mutate(
  identifier2016=paste0(iso3_code,"_2016"),
  identifier2020=paste0(iso3_code,"_2020"),
  P_mort_2020=`2020`,
  `2020`= NULL,
  P_mort_2021=`2021`,
  `2021`= NULL,
  `NA`<-NULL,
  Region<-NULL
  )  ->excessCounrty_year_pvalue3


# excessCounrty_year_pvalue3 %>% 
# fwrite("../output/02_ExcessMort_annual.csv")
```


```{r}
ExcessMort_annual<-fread("../output/02_ExcessMort_annual.csv")

ExcessMort_annual %>% dplyr::select(country,Region,`Income group`,pscore_mean,pscore_cat,iso3_code,P_mort_2020,P_mort_2021)-> ExcessMort_annual_sd


```


### 2. WDI Dimred
```{r,message=FALSE}
wdi_data_cons_df<-fread("../output/01_wdi_dimred_2010to2021_k250.csv",header = T)
thisdimensions<-wdi_data_cons_df[,c("Short Name","Country Code","year","Income Group","PC1","PC2","PC3","PC4","PC5","PC6","PC7","PC8","iso1","iso2","iso3","iso4","iso5","iso6","iso7","iso8")]

```
```{r}
require(scales)
require(dplyr)
wdi_data_cons_df %>% filter(year%in% c(2016:2022)) %>% dplyr::select(year,RegionSoviet,NY.GDP.PCAP.CD,iso1,iso2,iso3,iso4,iso5) %>% 
  group_by(year,RegionSoviet) %>% summarise(`GDP per capita`=mean(NY.GDP.PCAP.CD,na.rm = T),iso1=mean(iso1,na.rm = T)
                                            ,iso2=mean(iso2,na.rm = T)
                                            ,iso3=mean(iso3,na.rm = T)
                                            ,iso5=mean(iso5,na.rm = T)
                                            ,iso4=mean(iso4,na.rm = T)) %>% 
   pivot_longer(names_to = "var",
               "GDP per capita":"iso4",
               values_drop_na = TRUE) %>% 
  filter(var%in%c("GDP per capita","iso1","iso2","iso4"))%>% ggplot(aes(x=year,y=(value),color=RegionSoviet))+ geom_point() +geom_line() +theme_classic2() +ylab("")+ scale_x_continuous(breaks=c(2010,2012,2014,2016,2018,2020,2022))+theme(legend.position = "asd")+ scale_color_brewer(palette = "Set1")+ facet_wrap(~var,scales = "free_y")
ggsave("../figures/02_trajectories.png",device = "png",width = 9,height = 7)
wdi_data_cons_df %>% filter(year%in% c(2016:2022)) %>% dplyr::select(year,RegionSoviet,NY.GDP.PCAP.CD,iso1,iso2,iso3,iso4,iso5) %>% 
  group_by(year,RegionSoviet) %>% summarise(`GDP per capita`=mean(NY.GDP.PCAP.CD,na.rm = T),iso1=mean(iso1,na.rm = T)
                                            ,iso2=mean(iso2,na.rm = T)
                                            ,iso3=mean(iso3,na.rm = T)
                                            ,iso5=mean(iso5,na.rm = T)
                                            ,iso4=mean(iso4,na.rm = T)) %>% 
   pivot_longer(names_to = "var",
               "GDP per capita":"iso4",
               values_drop_na = TRUE) %>% 
  filter(var%in%c("GDP per capita","iso1","iso2","iso4"))%>% ggplot(aes(x=year,y=(value),color=RegionSoviet))+ geom_point() +geom_line() +theme_classic2() +ylab("")+ scale_x_continuous(breaks=c(2010,2012,2014,2016,2018,2020,2022))+theme(legend.position = "bottom")+ scale_color_brewer(palette = "Set1")+ facet_wrap(~var,scales = "free_y")

ggsave("../figures/02_trajectories_legend.png",device = "png",width = 9,height = 7)
```


### Combine. Merge
```{r,message=F}

dimred2021_knn250_pmort<-merge(wdi_data_cons_df,ExcessMort_annual_sd,by.x="Country Code",by.y="iso3_code",all.x = T) 
#dimred2021_knn250_pmort %>% fwrite("../output/02_230612_Dimred_ExcessMort_annual.csv")

```


# 1 Linear prediction {.tabset}


## Dimred19~P20 

```{r}

lm(scale(P_mort_2020)~scale(iso1)+
     scale(iso2)+
     scale(iso3)+
     scale(iso4)+
     scale(iso5)+
     scale(iso6)+
     scale(iso7)+
     scale(iso8)+
     scale(iso9)+
     scale(iso10),dimred2021_knn250_pmort[dimred2021_knn250_pmort$year%in%2019,]) %>%  summary


```




## Dimred19~P21
```{r}
lm(scale(P_mort_2021)~scale(iso1)+
     scale(iso2)+
     scale(iso3)+
     scale(iso4)+
     scale(iso5)+
     scale(iso6)+
     scale(iso7)+
     scale(iso8)+
     scale(iso9)+
     scale(iso10),dimred2021_knn250_pmort[dimred2021_knn250_pmort$year%in%2019,]) %>%  summary
```

## Dimred21~P20/21 mean



```{r}
lm(scale(pscore_mean)~scale(iso1)+
     scale(iso2)+
     scale(iso3)+
     scale(iso4)+
     scale(iso5)+
     scale(iso6)+
     scale(iso7)+
     scale(iso8)+
     scale(iso9)+
     scale(iso10),dimred2021_knn250_pmort[dimred2021_knn250_pmort$year%in%2019,]) %>%  summary 
```

## Dimred20~P20 

```{r}

lm(scale(P_mort_2020)~scale(iso1)+
     scale(iso2)+
     scale(iso3)+
     scale(iso4)+
     scale(iso5)+
     scale(iso6)+
     scale(iso7)+
     scale(iso8)+
     scale(iso9)+
     scale(iso10),dimred2021_knn250_pmort[dimred2021_knn250_pmort$year%in%2020,]) %>%  summary %>% tab_model(show.stat=T )


```

## Dimred20~P21
```{r}

lm(scale(P_mort_2021)~scale(iso1)+
     scale(iso2)+
     scale(iso3)+
     scale(iso4)+
     scale(iso5)+
     scale(iso6)+
     scale(iso7)+
     scale(iso8)+
     scale(iso9)+
     scale(iso10),dimred2021_knn250_pmort[dimred2021_knn250_pmort$year%in%2020,]) %>%  summary %>% tab_model(show.stat=T ) 

```

## Dimred21~P21 



```{r}
lm(scale(P_mort_2021)~scale(iso1)+
     scale(iso2)+
     scale(iso3)+
     scale(iso4)+
     scale(iso5)+
     scale(iso6)+
     scale(iso7)+
     scale(iso8)+
     scale(iso9)+
     scale(iso10),dimred2021_knn250_pmort[dimred2021_knn250_pmort$year%in%2021,]) %>%  summary %>% tab_model(show.stat=T ) 
```

# 1.5 Coefficient estimation

```{r}
dimred3<-dimred2021_knn250_pmort[dimred2021_knn250_pmort$year%in%2019,]  %>%  mutate(p_mort_text=case_when(P_mort_2020>40~`Country Code`,
                               P_mort_2020<(-5)~`Country Code`,
                               TRUE~"")) %>%  
  ggplot(aes(y=P_mort_2020,x=iso3,label=p_mort_text))+ geom_point(alpha=0.8)+ geom_hline(yintercept =0)+geom_text(aes(color=`Income Group`))+ geom_smooth(method = "lm") +   theme_tufte()+ xlab("Isomap Dimension 3") + scale_color_brewer(palette = "Dark2")+  ggtitle("2019 Isomap Dimension 3 ~ P Mort 2020") + theme(legend.position = "bottom") 

dimred4<-dimred2021_knn250_pmort[dimred2021_knn250_pmort$year%in%2019,]  %>%  mutate(p_mort_text=case_when(P_mort_2020>40~`Country Code`,
                               P_mort_2020<(-5)~`Country Code`,
                               TRUE~"")) %>%  
  ggplot(aes(y=P_mort_2020,x=iso4,label=p_mort_text))+ geom_point(alpha=0.8)+ geom_hline(yintercept =0)+geom_text(aes(color=`Income Group`))+ geom_smooth(method = "lm") +   theme_tufte()+ xlab("Isomap Dimension 4") + scale_color_brewer(palette = "Dark2")+  ggtitle("2019 Isomap Dimension 4 ~ P Mort 2020") + theme(legend.position = "bottom") 

suppressWarnings(ggarrange(dimred3,dimred4))
```

```{r}
dimred4<-dimred2021_knn250_pmort[dimred2021_knn250_pmort$year%in%2019,]  %>%  mutate(p_mort_text=case_when(P_mort_2020>40~`Country Code`,
                               P_mort_2020<(-5)~`Country Code`,
                               TRUE~"")) %>%  
  ggplot(aes(y=P_mort_2020,x=iso4,label=p_mort_text))+ geom_point(alpha=0.8)+ geom_hline(yintercept =0)+geom_text(aes(color=`Income Group`))+ geom_smooth(method = "lm") +   theme_tufte()+ xlab("Isomap Dimension 4") + scale_color_brewer(palette = "Dark2")+  ggtitle("2019 Isomap Dimension 4 ~ P Mort 2020") + theme(legend.position = "bottom") 
```


## Summary plots {.tabset}

```{r}
a<-lm(P_mort_2020~iso1+iso2+iso3+iso4+iso5+iso6+iso7+iso8+iso9+iso10,dimred2021_knn250_pmort[dimred2021_knn250_pmort$year%in%2019]) %>%  summary
b<-lm(P_mort_2020~iso1+iso2+iso3+iso4+iso5+iso6+iso7+iso8+iso9+iso10,dimred2021_knn250_pmort[dimred2021_knn250_pmort$year%in%2020]) %>%  summary
c<-lm(P_mort_2021~iso1+iso2+iso3+iso4+iso5+iso6+iso7+iso8+iso9+iso10,dimred2021_knn250_pmort[dimred2021_knn250_pmort$year%in%2020]) %>%  summary
d<-lm(P_mort_2021~iso1+iso2+iso3+iso4+iso5+iso6+iso7+iso8+iso9+iso10,dimred2021_knn250_pmort[dimred2021_knn250_pmort$year%in%2021]) %>%  summary
e<-lm(pscore_mean~iso1+iso2+iso3+iso4+iso5+iso6+iso7,dimred2021_knn250_pmort[dimred2021_knn250_pmort$year%in%2019]) %>%  summary
lm(pscore_mean~iso1+iso2+iso3+iso4+iso5+iso6+iso7,dimred2021_knn250_pmort[dimred2021_knn250_pmort$year%in%2019]) %>%  summary
```

s


```{r}
require(GGally)

aplot<-ggcoef_model(a) + ggtitle("WDI19~P20" )%>% suppressWarnings()
bplot<-ggcoef_model(b)+  ggtitle("WDI20~P20" )%>% suppressWarnings()

cplot<-ggcoef_model(c) + ggtitle("WDI20~P21" )%>% suppressWarnings()
dplot<-ggcoef_model(d)+  ggtitle("WDI21~P21" )%>% suppressWarnings()
```



### WDI~P20
```{r}
ggarrange(aplot,bplot)%>% suppressWarnings()

```
### WDI~P21
```{r}
ggarrange(cplot,dplot)%>% suppressWarnings()
```
### WDI20 ~ P_all_Covariates



#2 Clustering of the Isos

## 2019 Dimred {.tabset}
```{r}
wdi_data_cons_df<-dimred2021_knn250_pmort

isos2019_to2020<-wdi_data_cons_df[wdi_data_cons_df$year%in%2019,c("Country Code","pscore_mean","P_mort_2020","P_mort_2021","iso1","iso2","iso3","iso4","iso5","iso6","iso7","iso8","iso9","iso10")]

# Load the required library
library(apcluster)

# Perform Affinity Propagation clustering
isos2019_to2020 %>%   na.omit()->isos2019_to2020
  # Calculate the similarity matrix
cluster <- apcluster(negDistMat(r=2), scale(isos2019_to2020[,c("iso1","iso2","iso3","iso4")]) , q=0.01) 

# Explore the resulting clusters
# cluster_summary <- aggregate(data$x, by = list(affinity_result$cluster, affinity_result$label), FUN = mean)


exemplars <- integer(length(cluster@exemplars)) 

for(i in seq_along(cluster@exemplars)) {
      exemplars[i] <- cluster@exemplars[[i]][1]
    }

  
cluster_ids <- integer(cluster@l)
  for(i in seq_along(cluster@clusters)) {
    cl <- cluster@clusters[[i]]
    for(j in seq_along(cl)) {
      cluster_ids[cl[j]] <- i
    }
  }
  
  # glimpse(clustered)
isos2019_to2020$clusterid<-cluster_ids
  require(viridis)
death_2020<-isos2019_to2020 %>%
  pivot_longer(names_to = "isos",
               "iso1":"iso4",
               values_drop_na = TRUE) %>% 
  ggplot(aes(x=scale(value),y=P_mort_2020,color=as.factor(clusterid)))+ geom_point() +facet_wrap(~isos) +scale_color_brewer(palette = "Set1")+ggtitle("2020")+theme(legend.position = "asd")

death_2021<-isos2019_to2020 %>%
  pivot_longer(names_to = "isos",
               "iso1":"iso4",
               values_drop_na = TRUE) %>% 
  ggplot(aes(x=scale(value),y=P_mort_2021,color=as.factor(clusterid)))+ geom_point() +facet_wrap(~isos) +scale_color_brewer(palette = "Set1") +ggtitle("2021")+theme(legend.position = "asd")
death_2021
death_all<-isos2019_to2020 %>%
  pivot_longer(names_to = "isos",
               "iso1":"iso4",
               values_drop_na = TRUE) %>% 
  ggplot(aes(x=value,y=pscore_mean,color=as.factor(clusterid)))+ geom_point() +facet_wrap(~isos) +scale_color_brewer(palette = "Set1")+ggtitle("2020/2021")
```

### 2020
```{r}
death_2020
```

### 2021
```{r}
death_2021
```

### 2020/21
```{r}
death_all
```

## 2020 Dimred {.tabset}
```{r}
wdi_data_cons_df<-dimred2021_knn250_pmort

isos2019_to2020<-wdi_data_cons_df[wdi_data_cons_df$year%in%2020,c("Country Code","pscore_mean","P_mort_2020","P_mort_2021","iso1","iso2","iso3","iso4","iso5","iso6","iso7","iso8","iso9","iso10")]

# Load the required library
library(apcluster)

# Perform Affinity Propagation clustering
isos2019_to2020 %>%   na.omit()->isos2019_to2020
  # Calculate the similarity matrix
cluster <- apcluster(negDistMat(r=2), scale(isos2019_to2020[,c("iso1","iso2","iso3","iso4")]) , q=0.01) 

# Explore the resulting clusters
# cluster_summary <- aggregate(data$x, by = list(affinity_result$cluster, affinity_result$label), FUN = mean)


exemplars <- integer(length(cluster@exemplars)) 

for(i in seq_along(cluster@exemplars)) {
      exemplars[i] <- cluster@exemplars[[i]][1]
    }

  
cluster_ids <- integer(cluster@l)
  for(i in seq_along(cluster@clusters)) {
    cl <- cluster@clusters[[i]]
    for(j in seq_along(cl)) {
      cluster_ids[cl[j]] <- i
    }
  }
  
  # glimpse(clustered)
isos2019_to2020$clusterid<-cluster_ids
  require(viridis)
death_2020<-isos2019_to2020 %>%
  pivot_longer(names_to = "isos",
               "iso1":"iso4",
               values_drop_na = TRUE) %>% 
  ggplot(aes(x=value,y=P_mort_2020,color=as.factor(clusterid)))+ geom_point() +facet_wrap(~isos) +scale_color_brewer(palette = "Set1")+ggtitle("2020")+theme(legend.position = "asd")

death_2021<-isos2019_to2020 %>%
  pivot_longer(names_to = "isos",
               "iso1":"iso4",
               values_drop_na = TRUE) %>% 
  ggplot(aes(x=value,y=P_mort_2021,color=as.factor(clusterid)))+ geom_point() +facet_wrap(~isos) +scale_color_brewer(palette = "Set1") +ggtitle("2021")+theme(legend.position = "asd")

death_all<-isos2019_to2020 %>%
  pivot_longer(names_to = "isos",
               "iso1":"iso4",
               values_drop_na = TRUE) %>% 
  ggplot(aes(x=value,y=pscore_mean,color=as.factor(clusterid)))+ geom_point() +facet_wrap(~isos) +scale_color_brewer(palette = "Set1")+ggtitle("2020/2021")

```
### 2020
```{r}
death_2020
```

### 2021
```{r}
death_2021
```

### 2020/21
```{r}
death_all
```

## 2021 Dimred {.tabset}
```{r}
wdi_data_cons_df<-dimred2021_knn250_pmort

isos2019_to2020<-wdi_data_cons_df[wdi_data_cons_df$year%in%2021,c("Country Code","pscore_mean","P_mort_2020","P_mort_2021","iso1","iso2","iso3","iso4","iso5","iso6","iso7","iso8","iso9","iso10")]

# Load the required library
library(apcluster)

# Perform Affinity Propagation clustering
isos2019_to2020 %>%   na.omit()->isos2019_to2020
  # Calculate the similarity matrix
cluster <- apcluster(negDistMat(r=2), scale(isos2019_to2020[,c("iso1","iso2","iso3","iso4")]) , q=0.01) 

# Explore the resulting clusters
# cluster_summary <- aggregate(data$x, by = list(affinity_result$cluster, affinity_result$label), FUN = mean)


exemplars <- integer(length(cluster@exemplars)) 

for(i in seq_along(cluster@exemplars)) {
      exemplars[i] <- cluster@exemplars[[i]][1]
    }
cluster_ids <- integer(cluster@l)
  for(i in seq_along(cluster@clusters)) {
    cl <- cluster@clusters[[i]]
    for(j in seq_along(cl)) {
      cluster_ids[cl[j]] <- i
    }
  }
  
  # glimpse(clustered)
isos2019_to2020$clusterid<-cluster_ids
  require(viridis)
death_2020<-isos2019_to2020 %>%
  pivot_longer(names_to = "isos",
               "iso1":"iso4",
               values_drop_na = TRUE) %>% 
  ggplot(aes(x=value,y=P_mort_2020,color=as.factor(clusterid)))+ geom_point() +facet_wrap(~isos) +scale_color_brewer(palette = "Set1")+ggtitle("2020")+theme(legend.position = "asd")

death_2021<-isos2019_to2020 %>%
  pivot_longer(names_to = "isos",
               "iso1":"iso4",
               values_drop_na = TRUE) %>% 
  ggplot(aes(x=value,y=P_mort_2021,color=as.factor(clusterid)))+ geom_point() +facet_wrap(~isos) +scale_color_brewer(palette = "Set1") +ggtitle("2021")+theme(legend.position = "asd")

death_all<-isos2019_to2020 %>%
  pivot_longer(names_to = "isos",
               "iso1":"iso4",
               values_drop_na = TRUE) %>% 
  ggplot(aes(x=value,y=pscore_mean,color=as.factor(clusterid)))+ geom_point() +facet_wrap(~isos) +scale_color_brewer(palette = "Set1")+ggtitle("2020/2021")

```

### 2020
```{r}
death_2020
```

### 2021
```{r}
death_2021
```

### 2020/21
```{r}
death_all
```


## 2020/2021  Dimred + Pmort {.tabset}
```{r}

wdi_data_cons_df<-dimred2021_knn250_pmort

isos2019_to2021<-wdi_data_cons_df[wdi_data_cons_df$year%in%2021,c("Country Code","P_mort_2021","iso1","iso2","iso3","iso4","iso5","iso6","iso7","iso8","iso9","iso10","year","pscore_mean","RegionSoviet","Income Group")]
isos2019_to2020<-wdi_data_cons_df[wdi_data_cons_df$year%in%2020,c("Country Code","P_mort_2020","iso1","iso2","iso3","iso4","iso5","iso6","iso7","iso8","iso9","iso10","year","pscore_mean","RegionSoviet","Income Group")]

isos2019_to2020_2<-isos2019_to2020 %>% rename(Pmort=P_mort_2020)
isos2019_to2021_2<-isos2019_to2021 %>% rename(Pmort=P_mort_2021)

isos2020_2021<-rbind(isos2019_to2020_2,isos2019_to2021_2)
# Load the required library
library(apcluster)

# Perform Affinity Propagation clustering
isos2020_2021 %>%   na.omit()->isos2020_2021

  # Calculate the similarity matrix
set.seed(19) 
# cluster <- apcluster(negDistMat(r=2), scale(isos2020_2021[,c("Pmort","iso1","iso2","iso3","iso4","iso5","iso6")]) , q=0.0001) 
kmeans_result <- kmeans(scale(isos2020_2021[,c("iso1","iso2","iso3","iso4","iso5")]), centers = 6)
cluster_assignments <- kmeans_result$cluster
isos2020_2021$Kcluster <- cluster_assignments

?kmeans

isos2020_2021 %>% filter(year%in% c(2020,2021)) %>% 
  pivot_longer(names_to = "isos",
               "iso1":"iso4",
               values_drop_na = TRUE) %>% 
  ggplot(aes(x=scale(value),y=(pscore_mean),color=as.factor(Kcluster)))+ geom_point() +facet_wrap(~isos,scales = "free_x") +scale_color_brewer("Cluster",palette = "Dark2")+ggtitle("Excess Mortality in the latent space ")+ xlab("")+ theme_bw()+ theme(legend.position = "bottom",strip.text.x = element_text(size = 12, colour = "Black")) +ylab("P Score Excess Mortality")
#ggsave("../figures/02_ExcessMort_cluster.png",device = "png",width =9,height = 7,dpi = 320)

```
```{r}
ordera<-c("North America","Sub-Saharan Africa","East Asia & Pacific","Europe & Central Asia","Post-Soviet","Latin America & Caribbean","Middle East & North Africa", "South Asia") 

```

```{r}


sizes <- factor(as.factor(isos2020_2021$RegionSoviet), levels=ordera)
isos2020_2021$RegionSoviet2<-sizes
isos2020_2021 %>% filter(year%in% c(2020,2021)) %>% 
  pivot_longer(names_to = "isos",
               "iso1":"iso4",
               values_drop_na = TRUE) %>% 
  ggplot(aes(x=scale(value),y=(pscore_mean),color=as.factor(RegionSoviet2)))+ geom_point() +facet_wrap(~isos,scales = "free_x") +scale_color_brewer("Regions",palette = "Dark2")+ggtitle("Excess Mortality in the world")+ xlab("")+ylab("P Score Excess Mortality")+ theme_bw()+ theme(legend.position = "bottom") 

ggsave("../figures/02_ExcessMort_worldregions.png",device = "png",width =9,height = 7,dpi = 320)
```


```{r}
isos2020_2021 %>% filter(year%in% 2021) %>% 
  pivot_longer(names_to = "isos",
               "iso1":"iso4",
               values_drop_na = TRUE) %>% 
  ggplot(aes(x=scale(value),y=(pscore_mean),color=as.factor(`Income Group`)))+ geom_point() +facet_wrap(~isos,scales = "free_x") +scale_color_brewer("Regions",palette = "Dark2")+ggtitle("Excess Mortality in Income Groups")+ xlab("")+ylab("P Score Excess Mortality")+ theme_bw()+ theme(legend.position = "bottom") 
```




# 3 Find bet correlation method
```{r, warning=FALSE,message=FALSE}
wdi_data_cons_df2<-wdi_data_cons_df %>% data.frame()
library(minerva)
library(Hmisc)
variable<-"VC.IHR.PSRC.P5"


idx <- !is.na(wdi_data_cons_df2[,c('iso3',paste(variable))])
idx2<-idx %>% apply( . ,1, function(x)all(x)==T)
energy::dcor(wdi_data_cons_df2[idx2,variable],
                 wdi_data_cons_df2[idx2,'iso3'])

data1<-data.frame(wdi_data_cons_df2[idx2,variable]) %>%  unlist %>%  as.vector()
data2<-data.frame(wdi_data_cons_df2[idx2,'iso3']) %>%  unlist %>%  as.vector()

mutual_information <- entropy::entropy(data1,data2)

hoeffd(data1,
                data2)


mic_value <- minerva::mine(data1,data2)
mic_value

combined_data <- data.frame(wdi_data_cons_df2[idx2,variable],
                 wdi_data_cons_df2[idx2,'iso3'])



# Calculate the mutual information
wdi_data_cons_df[,c("Country Code","year","iso1")]
```
dCOR is the best

# 4  Features in Dimensions
```{r}
dcorTable_all<-fread("../data/01_CorrWDI_dimred_10.csv")

besti<-merge(wdi_explained[,c("Series Code","Topic","Indicator Name")],dcorTable_all,by.x = "Series Code",by.y="Indicators") 
besti %>% datatable
```





## Plot Topics{.tabset}

### ISO3
Third quantile
```{r}
besti %>% filter(iso3>round(besti$iso3 %>% quantile(),digits = 3)[3],
                 iso1<round(besti$iso1 %>% quantile(),digits = 3)[3],
                 
                 
                 ) %>% arrange(desc(iso3))->isoExtra
isoExtra %>%  pivot_longer(names_to = "isos",
               "iso1":"iso10",
               values_drop_na = TRUE)  %>% 
  mutate(labelA=str_extract(Topic,pattern = "^.[a-zA_Z].+[^:{1}]"),
         labelB=str_extract(Topic,pattern = "^.[a-zA_Z]+"),
         
         ) ->isoExtraplot
isoExtraplot$numericiso<-str_remove_all(isoExtraplot$isos,"iso")
isoExtraplot %>%  
  ggplot(data= . ,aes(reorder(x=isos,sort(as.numeric(numericiso))),y=value))+ geom_boxplot()+ geom_jitter(data=isoExtraplot[isoExtraplot$isos%in% c("iso1","iso2","iso3","iso4","iso10"),],aes(color=isos)) +facet_wrap(~labelB,ncol = 5)+ theme_bw()+ scale_color_brewer(palette = "Dark2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ xlab("")

#ggsave("../fig/230404_iso3.png",device = "png",dpi = 900)  

```
Social Protection & Labor: Labor force structure
Private Sector & Trade: Imports
Enviromental emissions
Enviromental landuse
Adjusted savings: carbon dioxide damage (% of GNI)

### ISO3 Subt
Third quantile
```{r}
besti %>% filter(iso3>round(besti$iso3 %>% quantile(),digits = 3)[3],
                 iso1<round(besti$iso1 %>% quantile(),digits = 3)[3],
                 
                 
                 ) %>% arrange(desc(iso3))->isoExtra
isoExtra %>%  pivot_longer(names_to = "isos",
               "iso1":"iso10",
               values_drop_na = TRUE)  %>% 
  mutate(labelA=str_extract(Topic,pattern = "^.[a-zA_Z].+[^:{1}]"),
         labelB=str_extract(Topic,pattern = "^.[a-zA_Z]+"),
         labelC=str_extract(Topic,pattern = "^.[a-zA_Z].+:")
         
         ) ->isoExtraplot

isoExtraplot$numericiso<-str_remove_all(isoExtraplot$isos,"iso")
isoExtraplot[isoExtraplot$isos%in% c("iso1","iso2","iso3","iso4","iso4","iso5") &grepl("Enviroment|Education|Policy & debt|Social Protection|Private|Public|Health:$",isoExtraplot$labelC),] %>%  
  ggplot(data= . ,aes(reorder(x=isos,sort(as.numeric(numericiso))),y=value,color=isos))+ geom_boxplot() +facet_wrap(~labelC,ncol = 3)+ theme_tufte()+ scale_color_brewer(palette = "Set1")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ xlab("") + ggtitle("WDI ~ Iso3 and Iso 4")

#ggsave("../fig/04_230509_extraiso3.png",device = "png",dpi = 900)  

```

### ISO4
Third quantile
```{r}
besti %>% filter(iso4>round(besti$iso4 %>% quantile(),digits = 3)[3],
                 iso1<round(besti$iso1 %>% quantile(),digits = 3)[3],
                 
                 
                 ) %>% arrange(desc(iso4))->isoExtra
isoExtra %>%  pivot_longer(names_to = "isos",
               "iso1":"iso10",
               values_drop_na = TRUE)  %>% 
  mutate(labelA=str_extract(Topic,pattern = "^.[a-zA_Z].+[^:{1}]"),
         labelB=str_extract(Topic,pattern = "^.[a-zA_Z]+"),
         
         ) ->isoExtraplot
isoExtraplot$numericiso<-str_remove_all(isoExtraplot$isos,"iso")
isoExtraplot %>%  
  ggplot(data= . ,aes(reorder(x=isos,sort(as.numeric(numericiso))),y=value))+ geom_boxplot()+ geom_jitter(data=isoExtraplot[isoExtraplot$isos%in% c("iso1","iso2","iso3","iso4","iso10"),],aes(color=isos)) +facet_wrap(~labelB,ncol = 5)+ theme_bw()+ scale_color_brewer(palette = "Dark2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ xlab("")

#ggsave("../fig/230404_iso4.png",device = "png",dpi = 900)  

```
Private Sector & Trade: Imports
Textiles and clothing (% of value added in manufacturing)
Public Sector: 

### ISO4 Subt
Third quantile
```{r}
besti %>% filter(iso3>round(besti$iso3 %>% quantile(),digits = 3)[3],
                 iso1<round(besti$iso1 %>% quantile(),digits = 3)[3],
                 
                 
                 ) %>% arrange(desc(iso4))->isoExtra

isoExtra %>%  pivot_longer(names_to = "isos",
               "iso1":"iso10",
               values_drop_na = TRUE)  %>% 
  mutate(labelA=str_extract(Topic,pattern = "^.[a-zA_Z].+[^:{1}]"),
         labelB=str_extract(Topic,pattern = "^.[a-zA_Z]+"),
         labelC=str_extract(Topic,pattern = "^.[a-zA_Z].+:")
         
         ) ->isoExtraplot

isoExtraplot$numericiso<-str_remove_all(isoExtraplot$isos,"iso")
isoExtraplot[isoExtraplot$isos%in% c("iso1","iso2","iso3","iso4","iso4","iso5")&grepl("Enviroment|Education|Policy & debt|Social Protection|Private|Public|Health:$",isoExtraplot$labelC), ,] %>%  
  ggplot(data= . ,aes(reorder(x=isos,sort(as.numeric(numericiso))),y=value,color=isos))+ geom_boxplot() +facet_wrap(~labelC,ncol = 3)+ theme_tufte()+ scale_color_brewer(palette = "Set1")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ xlab("") + ggtitle("WDI ~ Iso3")

#ggsave("../fig/04_230509_extraiso4.png",device = "png",dpi = 900)  

```

### ISO10
Third quantile
```{r}
besti %>% filter(iso4>round(besti$iso10 %>% quantile(),digits = 3)[3],
                 iso1<round(besti$iso1 %>% quantile(),digits = 3)[3],
                 
                 
                 ) %>% arrange(desc(iso10))->isoExtra
isoExtra %>%  pivot_longer(names_to = "isos",
               "iso1":"iso10",
               values_drop_na = TRUE)  %>% 
  mutate(labelA=str_extract(Topic,pattern = "^.[a-zA_Z].+[^:{1}]"),
         labelB=str_extract(Topic,pattern = "^.[a-zA_Z]+"),
         
         ) ->isoExtraplot
isoExtraplot$numericiso<-str_remove_all(isoExtraplot$isos,"iso")
isoExtraplot %>%  
  ggplot(data= . ,aes(reorder(x=isos,sort(as.numeric(numericiso))),y=value))+ geom_boxplot()+ geom_jitter(data=isoExtraplot[isoExtraplot$isos%in% c("iso1","iso2","iso3","iso4","iso10"),],aes(color=isos)) +facet_wrap(~labelB,ncol = 5)+ theme_bw()+ scale_color_brewer(palette = "Dark2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ xlab("")

#ggsave("../fig/230404_iso10.png",device = "png",dpi = 900)  

```


# 5 Latent space of other disasters

## Disastets


```{r}
# thidata<-fread("../data/230526_OWID_emdat/natural-disasters.csv")
# events<-colnames(disaster)[grepl("per 100,000",colnames(disaster))]
# disaster<-thidata[,!duplicated(colnames(thidata)),with=F]
# disaster$identifier<-paste0(disaster$`Country name`,"_",disaster$Year)
# dimred2021_knn250_pmort$identifier<-paste0(dimred2021_knn250_pmort$`Short Name`,"_",dimred2021_knn250_pmort$year)
```


```{r}
# 
# for(i in 1:length(events)){
#   
# 
# 
# disaster %>% filter(!is.na(get(events[i]))) %>% group_by(identifier) %>% 
# 
#   summarise(Disaster=mean(get(events[i]))) %>%  merge( . , dimred2021_knn250_pmort,by="identifier") %>%
# 
#   filter(!is.na(iso1),!is.na(Disaster))   ->thisone
#   if(nrow(thisone)>2){
#     a<-lm(scale(Disaster)~iso1+iso2+iso3+iso4+iso5+iso6,thisone )%>%  summary
#     aplot<-ggcoef_model(a)+ggtitle(events[i])   %>% suppressWarnings()
# 
#     b<-thisone %>%filter(year%in% c(2015:2022),(`Income Group`)!="") %>%  
#             pivot_longer(names_to = "isos",
#                "iso1":"iso4",
#                values_drop_na = TRUE)
#     b%>% 
#     ggplot(aes(x=scale(value),y=scale(Disaster),color=`Income Group`))+ geom_point() +facet_wrap(~isos) +scale_color_brewer(palette = "Set1")+ theme_classic() +ggtitle(events[i])-> b_plot
#     suppressWarnings(ggarrange(aplot,b_plot)) %>% ggsave(file = paste0("../figures/Dimred_Disasters/",i,"_dimplot.png",sep=""),device = "png",width = 14,height = 8)
# 
#   }
# 
# 
# 
# 
# 
# 
# }
# plotlist
```


# 6 Add democracy
## make data
```{r}

vdemdata<-fread("../data/CopyOfvdem_data_cons_df_75.csv")
dimred2021_knn250_pmort$identifier<-paste0(dimred2021_knn250_pmort$`Country Code`,dimred2021_knn250_pmort$year,sep="_")
vdemdata$identifier<-paste0(vdemdata$`Country Code`,vdemdata$year,sep="_")
coolsubset<-merge(dimred2021_knn250_pmort,vdemdata,by="identifier",suffixes = c(".wdi",".vdem"))
coolsubset %>% filter(year.wdi%in%2020) %>% ggplot(aes(x=iso2.vdem,y=P_mort_2021))+ geom_point()
coolsubset %>% filter(year.wdi%in%2020) %>% ggplot(aes(x=iso1.vdem,y=iso1.wdi))+ geom_point()
coolsubset %>% filter(year.wdi%in%2020) %>% ggplot(aes(x=iso2.vdem,y=iso1.wdi))+ geom_point()
coolsubset %>% filter(year.wdi%in%2020) %>% ggplot(aes(x=PC1.vdem,y=iso1.vdem))+ geom_point()+ theme_classic()
lm(P_mort_2021~iso1.wdi+iso2.wdi+iso3.wdi+iso4.wdi+PC1.vdem+PC2.vdem+iso3.vdem,coolsubset %>% filter(year.wdi%in%2019)) %>%  summary
```
```{r}
coolsubset %>%  filter(year.wdi%in% 2020) %>%  ggplot(aes(x=iso4.wdi))+ geom_density()
```

```{r}
coolsubset %>% ggplot(aes(x=iso1.vdem,y=iso2.vdem,color=P_mort_2020))+ geom_point() + scale_color_viridis() + theme_classic()

```
## cca

```{r}

coolsubset2<-coolsubset[,(grepl("year|Country Code.wdi|so[0-9]",colnames(coolsubset))),with=F] %>% na.omit()

vdem_iso<-coolsubset2[,(grepl("year|Country Code.wdi|so[0-9].vdem",colnames(coolsubset2))),with=F]%>% data.frame()%>% na.omit() %>% filter(year.wdi%in% 2020)

wdi_iso<-coolsubset2[,(grepl("year|Country Code.wdi|so[0-9].wdi",colnames(coolsubset2))),with=F] %>% data.frame()%>% na.omit() %>% filter(year.wdi%in% 2020)
rownames(vdem_iso)<-vdem_iso$Country.Code.wdi
rownames(wdi_iso)<-wdi_iso$Country.Code.wdi
vdem_iso$Country.Code.wdi<-NULL
vdem_iso$year.wdi<-NULL
vdem_iso$year.vdem<-NULL
wdi_iso$`Country.Code.wdi`<-NULL
wdi_iso$year.wdi<-NULL
wdi_iso$year.vdem<-NULL


thisdata<-data.frame(vdem_iso[,],wdi_iso)  %>% data.frame()


```


## iso plots

```{r}

coolsubset$pscore_cat2<-factor(coolsubset$pscore_cat,levels = c("<0", "0 to <5","5 to <10" , "10 to <20" ,"20 to <30" ,"30 to <40", "40 to <50",">=50" ))
```


```{r}
coolsubset %>% filter(year.wdi%in% c(2010:2022)) %>% dplyr::select(pscore_cat2,year.wdi,RegionSoviet.vdem,NY.GDP.PCAP.CD,
                                                        iso1.wdi,iso2.wdi,iso3.wdi,iso4.wdi,iso5.wdi,
                                                          iso1.vdem,iso2.vdem,iso3.vdem,iso4.vdem,iso5.vdem) %>% 
  group_by(year.wdi,pscore_cat2) %>% summarise(`GDP per capita`=mean(NY.GDP.PCAP.CD,na.rm = T),
                                            iso1.wdi=mean(iso1.wdi,na.rm = T),
                                            iso2.wdi=mean(iso2.wdi,na.rm = T),
                                            iso3.wdi=mean(iso3.wdi,na.rm = T),
                                            iso5.wdi=mean(iso5.wdi,na.rm = T),
                                            iso4.wdi=mean(iso4.wdi,na.rm = T),
                                            
                                            iso1.vdem=mean(iso1.vdem,na.rm = T),
                                            iso2.vdem=mean(iso2.vdem,na.rm = T),
                                            iso3.vdem=mean(iso3.vdem,na.rm = T),
                                            iso5.vdem=mean(iso5.vdem,na.rm = T),
                                            iso4.vdem=mean(iso4.vdem,na.rm = T))%>% 
   pivot_longer(names_to = "var",
               "GDP per capita":"iso4.vdem",
               values_drop_na = TRUE) %>% 
  filter(var%in%c("GDP per capita","iso1.wdi","iso1.vdem"))%>% ggplot(aes(x=year.wdi,y=(value),color=pscore_cat2))+ geom_point() +geom_line() +theme_classic2() +ylab("")+ scale_x_continuous(breaks=c(2010,2012,2014,2016,2018,2020,2022))+ scale_color_brewer(palette = "Set1")+ facet_wrap(~var,scales = "free_y")  + theme(axis.text.x = element_text(face="bold", color="black",
                           size=10, angle=45))
```

```{r}
coolsubset %>% filter(year.wdi%in% c(2010:2022)) %>% dplyr::select(pscore_cat,year.wdi,RegionSoviet.vdem,NY.GDP.PCAP.CD,
                                                        iso1.wdi,iso2.wdi,iso3.wdi,iso4.wdi,iso5.wdi,
                                                          iso1.vdem,iso2.vdem,iso3.vdem,iso4.vdem,iso5.vdem) %>% 
  group_by(year.wdi,pscore_cat) %>% summarise(`GDP per capita`=mean(NY.GDP.PCAP.CD,na.rm = T),
                                            iso1.wdi=mean(iso1.wdi,na.rm = T),
                                            iso2.wdi=mean(iso2.wdi,na.rm = T),
                                            iso3.wdi=mean(iso3.wdi,na.rm = T),
                                            iso5.wdi=mean(iso5.wdi,na.rm = T),
                                            iso4.wdi=mean(iso4.wdi,na.rm = T),
                                            
                                            iso1.vdem=mean(iso1.vdem,na.rm = T),
                                            iso2.vdem=mean(iso2.vdem,na.rm = T),
                                            iso3.vdem=mean(iso3.vdem,na.rm = T),
                                            iso5.vdem=mean(iso5.vdem,na.rm = T),
                                            iso4.vdem=mean(iso4.vdem,na.rm = T))%>% 
   pivot_longer(names_to = "var",
               "GDP per capita":"iso4.vdem",
               values_drop_na = TRUE) %>% 
  filter(var%in%c("GDP per capita","iso2.wdi","iso2.vdem"))%>% ggplot(aes(x=year.wdi,y=(value),color=pscore_cat))+ geom_point() +geom_line() +theme_classic2() +ylab("")+ scale_x_continuous(breaks=c(2010,2012,2014,2016,2018,2020,2022))+ scale_color_brewer(palette = "Set1")+ facet_wrap(~var,scales = "free_y")
```

```{r}

coolsubset %>% filter(year.wdi%in% c(2010:2022)) %>% dplyr::select(pscore_cat,year.wdi,RegionSoviet.vdem,NY.GDP.PCAP.CD,
                                                        iso1.wdi,iso2.wdi,iso3.wdi,iso4.wdi,iso5.wdi,
                                                          iso1.vdem,iso2.vdem,iso3.vdem,iso4.vdem,iso5.vdem) %>% 
  group_by(year.wdi,pscore_cat) %>% summarise(`GDP per capita`=mean(NY.GDP.PCAP.CD,na.rm = T),
                                            iso1.wdi=mean(iso1.wdi,na.rm = T),
                                            iso2.wdi=mean(iso2.wdi,na.rm = T),
                                            iso3.wdi=mean(iso3.wdi,na.rm = T),
                                            iso5.wdi=mean(iso5.wdi,na.rm = T),
                                            iso4.wdi=mean(iso4.wdi,na.rm = T),
                                            
                                            iso1.vdem=mean(iso1.vdem,na.rm = T),
                                            iso2.vdem=mean(iso2.vdem,na.rm = T),
                                            iso3.vdem=mean(iso3.vdem,na.rm = T),
                                            iso5.vdem=mean(iso5.vdem,na.rm = T),
                                            iso4.vdem=mean(iso4.vdem,na.rm = T))%>% 
   pivot_longer(names_to = "var",
               "GDP per capita":"iso4.vdem",
               values_drop_na = TRUE) %>% 
  filter(var%in%c("GDP per capita","iso4.wdi","iso1.vdem"))%>% ggplot(aes(x=year.wdi,y=(value),color=pscore_cat))+ geom_point() +geom_line() +theme_classic2() +ylab("")+ scale_x_continuous(breaks=c(2010,2012,2014,2016,2018,2020,2022))+ scale_color_brewer(palette = "Set1")+ facet_wrap(~var,scales = "free_y")
```


```{r}

coolsubset  %>% filter(year.wdi%in% c(2010:2022),`Country Code.wdi`%in% "PER")%>% dplyr::select(pscore_cat,year.wdi,RegionSoviet.vdem,NY.GDP.PCAP.CD,
                                                        iso1.wdi,iso2.wdi,iso3.wdi,iso4.wdi,iso5.wdi,
                                                          iso1.vdem,iso2.vdem,iso3.vdem,iso4.vdem,iso5.vdem) %>% 
  pivot_longer(names_to = "var",
               c("iso2.wdi","iso1.vdem"),
               values_drop_na = TRUE) %>% 
   ggplot(aes(x=year.wdi,y=value,color=var))+ geom_point()+geom_line()+ scale_color_brewer(palette = "Set1")+ theme_classic()
```


