---
title: "04_WDI_subset_correlation"
author: "Kolja Nenoff"
date: "2023-06-12"
output:
  rmdformats::readthedown:
    code_folding: hide
    self_contained: true
    css: custom.css 
    lightbox: true
    gallery: true
    highlight: tango  
    toc_depth: 5
---

We have described the dimension and their relationship to Excess Mort.
Now we are looing at the subsets and correlate them with the WDI.
The decision will be based on PDP and Shapley values.
We will look for subset for 2020 and 2021

# 0 Data and Methods
```{r,include=T,message=FALSE}
rm(list = ls())
require(xgboost)
require(caret)
require(pdp)
require(iml)
require(data.table)
require(tidyverse)
require(parallel)

```

```{r}
ExcessDimred<-fread("../output/02_230612_Dimred_ExcessMort_annual.csv")
allcorr<-fread("../data/01_CorrWDI_dimred_10.csv")
colnames(allcorr)[2:11]<-paste(colnames(allcorr)[2:11],"_all",sep="")
```


```{r}
# wdi_data_raw<-fread("../data/WDI_data_230609/WDI_CSV/WDIData.csv",header = T)
# wdi_data <- wdi_data_raw %>%
#   mutate(...66 = NULL,
#          V67 = NULL,
#          `Indicator Name` = NULL) %>%
#   pivot_longer(names_to = "year",
#                "1960":"2022",
#                values_drop_na = TRUE) %>%
#   mutate(year = as.integer(year)) %>%
#   pivot_wider(names_from = "Indicator Code")
# fwrite(wdi_data,"../data/04_WDI_data.csv")
wdi_data<-fread("../data/04_WDI_data.csv")
wdi_explained<-fread("../data/WDI_data_230609/WDI_CSV/WDISeries.csv")
```

# 1 Correlation 2020

```{r}
ExcessDimred2<-ExcessDimred%>%  select(`Country Code`,year,iso1,iso2,iso3,iso4,iso5,iso6,iso7,iso8,iso9,iso10,P_mort_2020,P_mort_2021) %>% 
  mutate(identifer=paste0(`Country Code`,"_",year),
                        "iso1"=unlist(scale(iso1)[,1]),
                        "iso2"=scale(iso2)[,1],
                        "iso3"=scale(iso3)[,1],
                        "iso4"=scale(iso4)[,1],
                        "iso5"=scale(iso5)[,1],
                        "iso6"=scale(iso6)[,1],
                        "iso7"=scale(iso7)[,1],
                        "iso8"=scale(iso8)[,1],
                        "iso9"=scale(iso9)[,1],
                        "iso10"=scale(iso10)[,1]) 

                  

wdi_data[,] %>%  mutate(identifer=paste0(`Country Code`,"_",year))-> wdi_data_2
# wer korreliert

corrWDItest<-right_join(wdi_data_2,ExcessDimred2[],by="identifer")
 
```

```{r,eval=TRUE}

# .cluster <- 10
#cl <- parallel::makeCluster(.cluster, outfile = "../data/04_corrLogfile.log")
# on.exit(parallel::stopCluster(cl))
wdi_data_cons_df<-corrWDItest %>% filter(!is.na(iso1))
used_var_names = wdi_explained$`Series Code`[wdi_explained$`Series Code` %in% colnames(corrWDItest)]


```




## iso 1 

```{r,eval=FALSE}
group_values<-wdi_data_cons_df$iso1 %>%  quantile()
thisiso<-paste0("iso1")
var_comb <- expand.grid(var_name = used_var_names,
                          iso_name = thisiso,
                          stringsAsFactors = FALSE)


wdi_data_cons_df %>%  mutate(categorie=case_when(
  iso1<group_values[2]~"0to25",
  iso1<group_values[3]~"25to50",
  iso1<group_values[4]~"50to75",
  iso1<group_values[5]~"75to100",
))-> wdi_data_cons_df

idx_0_25<-wdi_data_cons_df$categorie%in%"0to25"
idx_25_50<-wdi_data_cons_df$categorie%in%"25to50"
idx_50_75<-wdi_data_cons_df$categorie%in%"50to75"
idx_75_100<-wdi_data_cons_df$categorie%in%"75to100"

parallel::clusterEvalQ(cl, suppressWarnings(library(energy)))
parallel::clusterExport(cl, "wdi_data_cons_df", envir = environment())
parallel::clusterExport(cl, "var_comb", envir = environment())
parallel::clusterExport(cl, "idx_0_25", envir = environment())
parallel::clusterExport(cl, "idx_25_50", envir = environment())
parallel::clusterExport(cl, "idx_50_75", envir = environment())
parallel::clusterExport(cl, "idx_75_100", envir = environment())
```


```{r,eval=FALSE}
for(groupA in c("idx_0_25","idx_25_50","idx_50_75","idx_75_100")){
  

variable<-groupA

parallel::clusterExport(cl, "variable", envir = environment())
res <- parLapplyLB(cl, 1:nrow(var_comb), function (i) {
  message("i: ", i, " / ", nrow(var_comb))

  idx <- !is.na(wdi_data_cons_df[[var_comb$var_name[i]]])
  idx<-eval(parse(text=paste0("idx&",variable,sep="")))
  energy::dcor(wdi_data_cons_df[[var_comb$var_name[i]]][idx],
               wdi_data_cons_df[[var_comb$iso_name[i]]][idx])
  
})

res_Tsamples <- parLapplyLB(cl, 1:nrow(var_comb), function (i) {
  message("i: ", i, " / ", nrow(var_comb))
  idx <- !is.na(wdi_data_cons_df[[var_comb$var_name[i]]])
   idx<-eval(parse(text=paste0("idx&",variable,sep="")))
  sum(idx)
})
res_spear <- parLapplyLB(cl, 1:nrow(var_comb), function (i) {
  message("i: ", i, " / ", nrow(var_comb))
  idx <- !is.na(wdi_data_cons_df[[var_comb$var_name[i]]])
   idx<-eval(parse(text=paste0("idx&",variable,sep="")))
  cor(wdi_data_cons_df[[var_comb$var_name[i]]][idx],
               wdi_data_cons_df[[var_comb$iso_name[i]]][idx],method = "pearson")
})

wdi_data_cons_dcor <- matrix(
  NA_real_, length(used_var_names), 1,
  dimnames = list(used_var_names, thisiso)
)
for (i in seq_len(nrow(var_comb))) {
  wdi_data_cons_dcor[var_comb$var_name[i],
                     thisiso] <- res[[i]]
}

wdi_data_cons_dcor2<- matrix(
  NA_real_, length(used_var_names), 1,
  dimnames = list(used_var_names, thisiso)
)
for (i in seq_len(nrow(var_comb))) {
  wdi_data_cons_dcor2[var_comb$var_name[i],
                       thisiso] <- res_spear[[i]]
}
wdi_data_cons_dcor3<- matrix(
  NA_real_, length(used_var_names), 1,
  dimnames = list(used_var_names, thisiso)
)
for (i in seq_len(nrow(var_comb))) {
  wdi_data_cons_dcor3[var_comb$var_name[i],
                       thisiso] <- res_Tsamples[[i]]
}
colnames(wdi_data_cons_dcor2)<-paste("Pearson",colnames(wdi_data_cons_dcor2),sep="_")
colnames(wdi_data_cons_dcor3)<-paste("N",colnames(wdi_data_cons_dcor3),sep="_")

cbind(data.frame(wdi_data_cons_dcor),data.frame(N=wdi_data_cons_dcor3[,1]),wdi_data_cons_dcor2) %>% rownames_to_column(var = "indicators")-> outi
categoriestr<-str_replace(str_remove(groupA,"idx_"),"_","to")





outi$range<- paste0(round(min(wdi_data_cons_df[wdi_data_cons_df$categorie%in%categoriestr,..thisiso]),digits = 3)," to ",
       round(max(wdi_data_cons_df[wdi_data_cons_df$categorie%in%categoriestr,..thisiso]),digits = 3),sep="")
colnames(outi)<-paste(colnames(outi),categoriestr,sep = "_")

eval(parse(text=paste("outi-> WDI_dimred_iso1_",groupA,sep="")))

}

```
### save
```{r,eval=FALSE}
wdi_dimred_iso1<-cbind(WDI_dimred_iso1_idx_0_25,WDI_dimred_iso1_idx_25_50,WDI_dimred_iso1_idx_50_75,WDI_dimred_iso1_idx_75_100)
wdi_dimred_iso1_1<-data.frame("Indicators"=wdi_dimred_iso1$indicators_0to25,wdi_dimred_iso1[,!grepl("indicators",colnames(wdi_dimred_iso1))])



wdi_dimred_iso1_2<-merge(allcorr,wdi_dimred_iso1_1,by="Indicators",all.y=T)
wdi_dimred_iso1_3<-merge(wdi_explained[,c("Series Code","Topic","Indicator Name")],wdi_dimred_iso1_2,by.x="Series Code",by.y="Indicators")
#fwrite(wdi_dimred_iso1_3,"../results/corrTable/04_CorrTable_iso1_quantiles_all.csv")
```


```{r}
iso1_corr<-fread("../results/corrTable/04_CorrTable_iso1_quantiles_all.csv")

```



### subset -1 to 1 {.tabset}

#### age
```{r}
seriescode<-"SP.POP.0014.MA.ZS"

wdi_data_cons_df %>%  mutate(categorie=case_when(scale(iso1)>0.5~"Higher",
                                                 scale(iso1)<  -0.5~"Lower",
                                                  TRUE~"Iso1_peak"
                                                  ))%>% 
  filter(!is.na(get(seriescode)),!is.na(categorie)) -> yo
yo$categorie<-factor(as.factor(yo$categorie),levels=c("Lower","Iso1_peak","Higher"))
yo2<-yo[,colnames(yo)[colnames(yo)%in% c(paste(seriescode),"categorie")],with=F]
colnames(yo2)[1]<-"value"
series_label<-wdi_explained[wdi_explained$`Series Code` %in% seriescode,"Indicator Name"]
?ggbetweenstats
ggstatsplot::ggbetweenstats(,plot.type = "boxviolin",
                            p.adjust.method = "BH",
        point.args = list(alpha = 0),
      data = yo2,
      x = categorie,
      y = value ) + 
      # Add labels and title
      labs(x="",
           y = seriescode,
           title = series_label
      )+theme(
        # This is the new default font in the plot
        text = element_text(family = "Roboto", size = 10, color = "black"),
        plot.title = element_text(
          family = "Lobster Two", 
          size = 14,
          face = "bold",
          color = "#2a475e"
        ),
        # # Statistical annotations below the main title
        plot.subtitle = element_text(
          family = "Roboto",
          face = "bold",
          color="#1b2838"
        ),
        plot.title.position = "plot", # slightly different from default
        axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 12)
      )

```
#### age 2


```{r}
seriescode<-"SE.SEC.NENR"

wdi_data_cons_df %>%  mutate(categorie=case_when(scale(iso1)>0.5~"Higher",
                                                 scale(iso1)<  -0.5~"Lower",
                                                  TRUE~"Iso1_peak"
                                                  ))%>% 
  filter(!is.na(get(seriescode)),!is.na(categorie)) -> yo
yo$categorie<-factor(as.factor(yo$categorie),levels=c("Lower","Iso1_peak","Higher"))
yo2<-yo[,colnames(yo)[colnames(yo)%in% c(paste(seriescode),"categorie")],with=F]
colnames(yo2)[1]<-"value"
series_label<-wdi_explained[wdi_explained$`Series Code` %in% seriescode,"Indicator Name"]

ggstatsplot::ggbetweenstats(
      data = yo2,
      x = categorie,
      y = value ) + 
      # Add labels and title
      labs(x="",
           y = seriescode,
           title = series_label
      )+theme(
        # This is the new default font in the plot
        text = element_text(family = "Roboto", size = 10, color = "black"),
        plot.title = element_text(
          family = "Lobster Two", 
          size = 14,
          face = "bold",
          color = "#2a475e"
        ),
        # # Statistical annotations below the main title
        plot.subtitle = element_text(
          family = "Roboto",
          face = "bold",
          color="#1b2838"
        ),
        plot.title.position = "plot", # slightly different from default
        axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 12)
      )
```

#### Hospital beds 


```{r}
seriescode<-"SH.MED.BEDS.ZS"

wdi_data_cons_df %>%  mutate(categorie=case_when(scale(iso1)>0.5~"Higher",
                                                 scale(iso1)<  -0.5~"Lower",
                                                  TRUE~"Iso1_peak"
                                                  ))%>% 
  filter(!is.na(get(seriescode)),!is.na(categorie)) -> yo
yo$categorie<-factor(as.factor(yo$categorie),levels=c("Lower","Iso1_peak","Higher"))
yo2<-yo[,colnames(yo)[colnames(yo)%in% c(paste(seriescode),"categorie")],with=F]
colnames(yo2)[1]<-"value"
series_label<-wdi_explained[wdi_explained$`Series Code` %in% seriescode,"Indicator Name"]

ggstatsplot::ggbetweenstats(
      data = yo2,
      x = categorie,
      y = value ) + 
      # Add labels and title
      labs(x="",
           y = seriescode,
           title = series_label
      )+theme(
        # This is the new default font in the plot
        text = element_text(family = "Roboto", size = 10, color = "black"),
        plot.title = element_text(
          family = "Lobster Two", 
          size = 14,
          face = "bold",
          color = "#2a475e"
        ),
        # # Statistical annotations below the main title
        plot.subtitle = element_text(
          family = "Roboto",
          face = "bold",
          color="#1b2838"
        ),
        plot.title.position = "plot", # slightly different from default
        axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 12)
      )


ggsave("../figures/04_iso1_sig_hospitalbeds_boxpplot.png",device = "png")
```
#### Employment
```{r}

seriescode<-"SL.EMP.WORK.FE.ZS"

wdi_data_cons_df %>%  mutate(categorie=case_when(scale(iso1)>0.5~"Higher",
                                                 scale(iso1)<  -0.5~"Lower",
                                                  TRUE~"Iso1_peak"
                                                  ))%>% 
  filter(!is.na(get(seriescode)),!is.na(categorie)) -> yo
yo$categorie<-factor(as.factor(yo$categorie),levels=c("Lower","Iso1_peak","Higher"))
yo2<-yo[,colnames(yo)[colnames(yo)%in% c(paste(seriescode),"categorie")],with=F]
colnames(yo2)[1]<-"value"
series_label<-wdi_explained[wdi_explained$`Series Code` %in% seriescode,"Indicator Name"]

ggstatsplot::ggbetweenstats(
      data = yo2,
      x = categorie,
      y = value ) + 
      # Add labels and title
      labs(x="",
           y = seriescode,
           title = series_label
      )+theme(
        # This is the new default font in the plot
        text = element_text(family = "Roboto", size = 10, color = "black"),
        plot.title = element_text(
          family = "Lobster Two", 
          size = 14,
          face = "bold",
          color = "#2a475e"
        ),
        # # Statistical annotations below the main title
        plot.subtitle = element_text(
          family = "Roboto",
          face = "bold",
          color="#1b2838"
        ),
        plot.title.position = "plot", # slightly different from default
        axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 12)
      )
ggsave("../figures/04_iso1_sig_employment_boxpplot.png",device = "png")
```


## iso 2

```{r,eval=FALSE}
group_values<-wdi_data_cons_df$iso2 %>%  quantile()
thisiso<-paste0("iso2")
var_comb <- expand.grid(var_name = used_var_names,
                          iso_name = thisiso,
                          stringsAsFactors = FALSE)


wdi_data_cons_df %>%  mutate(categorie=case_when(
  iso2<group_values[2]~"0to25",
  iso2<group_values[3]~"25to50",
  iso2<group_values[4]~"50to75",
  iso2<group_values[5]~"75to100",
))-> wdi_data_cons_df

idx_0_25<-wdi_data_cons_df$categorie%in%"0to25"
idx_25_50<-wdi_data_cons_df$categorie%in%"25to50"
idx_50_75<-wdi_data_cons_df$categorie%in%"50to75"
idx_75_100<-wdi_data_cons_df$categorie%in%"75to100"

parallel::clusterEvalQ(cl, suppressWarnings(library(energy)))
parallel::clusterExport(cl, "wdi_data_cons_df", envir = environment())
parallel::clusterExport(cl, "var_comb", envir = environment())
parallel::clusterExport(cl, "idx_0_25", envir = environment())
parallel::clusterExport(cl, "idx_25_50", envir = environment())
parallel::clusterExport(cl, "idx_50_75", envir = environment())
parallel::clusterExport(cl, "idx_75_100", envir = environment())
```


```{r,eval=F}
for(groupA in c("idx_0_25","idx_25_50","idx_50_75","idx_75_100")){
  

variable<-groupA

parallel::clusterExport(cl, "variable", envir = environment())
res <- parLapplyLB(cl, 1:nrow(var_comb), function (i) {
  message("i: ", i, " / ", nrow(var_comb))

  idx <- !is.na(wdi_data_cons_df[[var_comb$var_name[i]]])
  idx<-eval(parse(text=paste0("idx&",variable,sep="")))
  energy::dcor(wdi_data_cons_df[[var_comb$var_name[i]]][idx],
               wdi_data_cons_df[[var_comb$iso_name[i]]][idx])
  
})

res_Tsamples <- parLapplyLB(cl, 1:nrow(var_comb), function (i) {
  message("i: ", i, " / ", nrow(var_comb))
  idx <- !is.na(wdi_data_cons_df[[var_comb$var_name[i]]])
   idx<-eval(parse(text=paste0("idx&",variable,sep="")))
  sum(idx)
})
res_spear <- parLapplyLB(cl, 1:nrow(var_comb), function (i) {
  message("i: ", i, " / ", nrow(var_comb))
  idx <- !is.na(wdi_data_cons_df[[var_comb$var_name[i]]])
   idx<-eval(parse(text=paste0("idx&",variable,sep="")))
  cor(wdi_data_cons_df[[var_comb$var_name[i]]][idx],
               wdi_data_cons_df[[var_comb$iso_name[i]]][idx],method = "pearson")
})

wdi_data_cons_dcor <- matrix(
  NA_real_, length(used_var_names), 1,
  dimnames = list(used_var_names, thisiso)
)
for (i in seq_len(nrow(var_comb))) {
  wdi_data_cons_dcor[var_comb$var_name[i],
                     thisiso] <- res[[i]]
}

wdi_data_cons_dcor2<- matrix(
  NA_real_, length(used_var_names), 1,
  dimnames = list(used_var_names, thisiso)
)
for (i in seq_len(nrow(var_comb))) {
  wdi_data_cons_dcor2[var_comb$var_name[i],
                       thisiso] <- res_spear[[i]]
}
wdi_data_cons_dcor3<- matrix(
  NA_real_, length(used_var_names), 1,
  dimnames = list(used_var_names, thisiso)
)
for (i in seq_len(nrow(var_comb))) {
  wdi_data_cons_dcor3[var_comb$var_name[i],
                       thisiso] <- res_Tsamples[[i]]
}
colnames(wdi_data_cons_dcor2)<-paste("Pearson",colnames(wdi_data_cons_dcor2),sep="_")
colnames(wdi_data_cons_dcor3)<-paste("N",colnames(wdi_data_cons_dcor3),sep="_")

cbind(data.frame(wdi_data_cons_dcor),data.frame(N=wdi_data_cons_dcor3[,1]),wdi_data_cons_dcor2) %>% rownames_to_column(var = "indicators")-> outi
categoriestr<-str_replace(str_remove(groupA,"idx_"),"_","to")





outi$range<- paste0(round(min(wdi_data_cons_df[wdi_data_cons_df$categorie%in%categoriestr,..thisiso]),digits = 3)," to ",
       round(max(wdi_data_cons_df[wdi_data_cons_df$categorie%in%categoriestr,..thisiso]),digits = 3),sep="")
colnames(outi)<-paste(colnames(outi),categoriestr,sep = "_")

eval(parse(text=paste("outi-> WDI_dimred_iso2_",groupA,sep="")))

}

```
### save
```{r,eval=FALSE}
wdi_dimred_iso2<-cbind(WDI_dimred_iso2_idx_0_25,WDI_dimred_iso2_idx_25_50,WDI_dimred_iso2_idx_50_75,WDI_dimred_iso2_idx_75_100)
wdi_dimred_iso2_1<-data.frame("Indicators"=wdi_dimred_iso2$indicators_0to25,wdi_dimred_iso2[,!grepl("indicators",colnames(wdi_dimred_iso2))])



wdi_dimred_iso2_2<-merge(allcorr,wdi_dimred_iso2_1,by="Indicators",all.y=T)
wdi_dimred_iso2_3<-merge(wdi_explained[,c("Series Code","Topic","Indicator Name")],wdi_dimred_iso2_2,by.x="Series Code",by.y="Indicators")
#fwrite(wdi_dimred_iso2_3,"../results/corrTable/04_CorrTable_iso2_quantiles_all.csv")
```

```{r}
wdi_dimred_iso2<-fread("../results/corrTable/04_CorrTable_iso2_quantiles_all.csv")
```
### age structure
```{r}
seriescode<-"SP.POP.7074.FE.5Y"
wdi_data_cons_df %>% filter(scale(iso1)<2&scale(iso1)>1.5,P_mort_2020>25) 
```


```{r}
wdi_data_cons_df %>%  mutate(categorie=case_when(scale(iso2)>1.5~"Iso2 peak  high",
                                                 scale(iso2)<  -0.5~"Iso2 peak  Low",
                                                 (scale(iso2)> (-0.5) &scale(iso2)<1.5)~"normal",
                                                  TRUE~"a"
                                                  ))%>% 
  filter(!is.na(get(seriescode)),!is.na(categorie)) -> yo
yo$categorie<-factor(as.factor(yo$categorie),levels=c("Iso2 peak  Low","normal","Iso2 peak  high","a"))
yo2<-yo[,colnames(yo)[colnames(yo)%in% c(paste(seriescode),"categorie")],with=F]
colnames(yo2)[1]<-"value"
series_label<-wdi_explained[wdi_explained$`Series Code` %in% seriescode,"Indicator Name"]

ggstatsplot::ggbetweenstats(
  plot.type = "boxviolin",
                            p.adjust.method = "BH",
        point.args = list(alpha = 0),

      data = yo2,
      x = categorie,
      y = value ) + 
      # Add labels and title
      labs(x="",
           y = seriescode,
           title = series_label
      )+theme(
        # This is the new default font in the plot
        text = element_text(family = "Roboto", size = 10, color = "black"),
        plot.title = element_text(
          family = "Lobster Two", 
          size = 14,
          face = "bold",
          color = "#2a475e"
        ),
        # # Statistical annotations below the main title
        plot.subtitle = element_text(
          family = "Roboto",
          face = "bold",
          color="#1b2838"
        ),
        plot.title.position = "plot", # slightly different from default
        axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 12)
      )


#ggsave("../figures/04_iso2_sig_agestructure_boxpplot.png",device = "png")
```
#### vulnarable work
```{r}
seriescode<-"SL.EMP.WORK.ZS"

wdi_data_cons_df %>% filter(year%in% c(2015:2021))%>%  mutate(categorie=case_when(scale(iso2)>1.5~"Iso2 peak  high",
                                                 scale(iso2)<  -0.5~"Iso2 peak  Low",
                                                  TRUE~"Iso2_normal"
                                                  ))%>% 
  filter(!is.na(get(seriescode)),!is.na(categorie)) -> yo
yo$categorie<-factor(as.factor(yo$categorie),levels=c("Iso2 peak  Low","Iso2_normal","Iso2 peak  high"))
yo2<-yo[,colnames(yo)[colnames(yo)%in% c(paste(seriescode),"categorie")],with=F]
colnames(yo2)[1]<-"value"
series_label<-wdi_explained[wdi_explained$`Series Code` %in% seriescode,"Indicator Name"]

ggstatsplot::ggbetweenstats(
      data = yo2,
      x = categorie,
      y = value ) + 
      # Add labels and title
      labs(x="",
           y = seriescode,
           title = series_label
      )+theme(
        # This is the new default font in the plot
        text = element_text(family = "Roboto", size = 10, color = "black"),
        plot.title = element_text(
          family = "Lobster Two", 
          size = 14,
          face = "bold",
          color = "#2a475e"
        ),
        # # Statistical annotations below the main title
        plot.subtitle = element_text(
          family = "Roboto",
          face = "bold",
          color="#1b2838"
        ),
        plot.title.position = "plot", # slightly different from default
        axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 12)
      )


#ggsave("../figures/04_iso2_sig_work_boxpplot.png",device = "png")
```
`



## iso 3 

```{r,eval=FALSE}
group_values<-wdi_data_cons_df$iso3 %>%  quantile()
thisiso<-paste0("iso3")
var_comb <- expand.grid(var_name = used_var_names,
                          iso_name = thisiso,
                          stringsAsFactors = FALSE)


wdi_data_cons_df %>%  mutate(categorie=case_when(
  iso3<group_values[2]~"0to25",
  iso3<group_values[3]~"25to50",
  iso3<group_values[4]~"50to75",
  iso3<group_values[5]~"75to100",
))-> wdi_data_cons_df

idx_0_25<-wdi_data_cons_df$categorie%in%"0to25"
idx_25_50<-wdi_data_cons_df$categorie%in%"25to50"
idx_50_75<-wdi_data_cons_df$categorie%in%"50to75"
idx_75_100<-wdi_data_cons_df$categorie%in%"75to100"

parallel::clusterEvalQ(cl, suppressWarnings(library(energy)))
parallel::clusterExport(cl, "wdi_data_cons_df", envir = environment())
parallel::clusterExport(cl, "var_comb", envir = environment())
parallel::clusterExport(cl, "idx_0_25", envir = environment())
parallel::clusterExport(cl, "idx_25_50", envir = environment())
parallel::clusterExport(cl, "idx_50_75", envir = environment())
parallel::clusterExport(cl, "idx_75_100", envir = environment())
```


```{r,eval=F}
for(groupA in c("idx_0_25","idx_25_50","idx_50_75","idx_75_100")){
  

variable<-groupA

parallel::clusterExport(cl, "variable", envir = environment())
res <- parLapplyLB(cl, 1:nrow(var_comb), function (i) {
  message("i: ", i, " / ", nrow(var_comb))

  idx <- !is.na(wdi_data_cons_df[[var_comb$var_name[i]]])
  idx<-eval(parse(text=paste0("idx&",variable,sep="")))
  energy::dcor(wdi_data_cons_df[[var_comb$var_name[i]]][idx],
               wdi_data_cons_df[[var_comb$iso_name[i]]][idx])
  
})

res_Tsamples <- parLapplyLB(cl, 1:nrow(var_comb), function (i) {
  message("i: ", i, " / ", nrow(var_comb))
  idx <- !is.na(wdi_data_cons_df[[var_comb$var_name[i]]])
   idx<-eval(parse(text=paste0("idx&",variable,sep="")))
  sum(idx)
})
res_spear <- parLapplyLB(cl, 1:nrow(var_comb), function (i) {
  message("i: ", i, " / ", nrow(var_comb))
  idx <- !is.na(wdi_data_cons_df[[var_comb$var_name[i]]])
   idx<-eval(parse(text=paste0("idx&",variable,sep="")))
  cor(wdi_data_cons_df[[var_comb$var_name[i]]][idx],
               wdi_data_cons_df[[var_comb$iso_name[i]]][idx],method = "pearson")
})

wdi_data_cons_dcor <- matrix(
  NA_real_, length(used_var_names), 1,
  dimnames = list(used_var_names, thisiso)
)
for (i in seq_len(nrow(var_comb))) {
  wdi_data_cons_dcor[var_comb$var_name[i],
                     thisiso] <- res[[i]]
}

wdi_data_cons_dcor2<- matrix(
  NA_real_, length(used_var_names), 1,
  dimnames = list(used_var_names, thisiso)
)
for (i in seq_len(nrow(var_comb))) {
  wdi_data_cons_dcor2[var_comb$var_name[i],
                       thisiso] <- res_spear[[i]]
}
wdi_data_cons_dcor3<- matrix(
  NA_real_, length(used_var_names), 1,
  dimnames = list(used_var_names, thisiso)
)
for (i in seq_len(nrow(var_comb))) {
  wdi_data_cons_dcor3[var_comb$var_name[i],
                       thisiso] <- res_Tsamples[[i]]
}
colnames(wdi_data_cons_dcor2)<-paste("Pearson",colnames(wdi_data_cons_dcor2),sep="_")
colnames(wdi_data_cons_dcor3)<-paste("N",colnames(wdi_data_cons_dcor3),sep="_")

cbind(data.frame(wdi_data_cons_dcor),data.frame(N=wdi_data_cons_dcor3[,1]),wdi_data_cons_dcor2) %>% rownames_to_column(var = "indicators")-> outi
categoriestr<-str_replace(str_remove(groupA,"idx_"),"_","to")





outi$range<- paste0(round(min(wdi_data_cons_df[wdi_data_cons_df$categorie%in%categoriestr,..thisiso]),digits = 3)," to ",
       round(max(wdi_data_cons_df[wdi_data_cons_df$categorie%in%categoriestr,..thisiso]),digits = 3),sep="")
colnames(outi)<-paste(colnames(outi),categoriestr,sep = "_")

eval(parse(text=paste("outi-> WDI_dimred_iso3_",groupA,sep="")))

}

```
### save
```{r,eval=FALSE}
wdi_dimred_iso3<-cbind(WDI_dimred_iso3_idx_0_25,WDI_dimred_iso3_idx_25_50,WDI_dimred_iso3_idx_50_75,WDI_dimred_iso3_idx_75_100)
wdi_dimred_iso3_1<-data.frame("Indicators"=wdi_dimred_iso3$indicators_0to25,wdi_dimred_iso3[,!grepl("indicators",colnames(wdi_dimred_iso3))])

colnames(allcorr)[2:11]<-paste(colnames(allcorr)[2:11],"_all",sep="")

wdi_dimred_iso3_2<-merge(allcorr,wdi_dimred_iso3_1,by="Indicators",all.y=T)
wdi_dimred_iso3_3<-merge(wdi_explained[,c("Series Code","Topic","Indicator Name")],wdi_dimred_iso3_2,by.x="Series Code",by.y="Indicators")
#fwrite(wdi_dimred_iso3_3,"../results/corrTable/04_CorrTable_iso3_quantiles_all.csv")
```

```{r}
iso3_corrtable<-fread("../results/corrTable/04_CorrTable_iso3_quantiles_all.csv")

```

#### share of textile production
```{r}
seriescode<-"NV.MNF.TXTL.ZS.UN"


wdi_data_cons_df %>%  mutate(categorie=case_when(scale(iso3)>1~"Iso3_negative",
                                                 scale(iso3)<  -0.5~"Iso3 peak  Low",
                                                  TRUE~"high_mort_no_peak"
                                                  ))%>% 
  filter(!is.na(get(seriescode)),!is.na(categorie)) -> yo

yo$categorie<-factor(as.factor(yo$categorie),levels=c("Iso3 peak  Low","high_mort_no_peak","Iso3_negative"))
yo2<-yo[,colnames(yo)[colnames(yo)%in% c(paste(seriescode),"categorie")],with=F]
colnames(yo2)[1]<-"value"
series_label<-wdi_explained[wdi_explained$`Series Code` %in% seriescode,"Indicator Name"]

ggstatsplot::ggbetweenstats(
      data = yo2,
      x = categorie,
      y = value ) + 
      # Add labels and title
      labs(x="",
           y = seriescode,
           title = series_label
      )+theme(
        # This is the new default font in the plot
        text = element_text(family = "Roboto", size = 10, color = "black"),
        plot.title = element_text(
          family = "Lobster Two", 
          size = 14,
          face = "bold",
          color = "#2a475e"
        ),
        # # Statistical annotations below the main title
        plot.subtitle = element_text(
          family = "Roboto",
          face = "bold",
          color="#1b2838"
        ),
        plot.title.position = "plot", # slightly different from default
        axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 12)
      )


ggsave("../figures/04_iso3_sig_textile_production_boxpplot.png",device = "png")
```
#### Unemployment
```{r}
seriescode<-"SL.UEM.TOTL.FE.ZS"

wdi_data_cons_df %>%  mutate(categorie=case_when(scale(iso3)>1~"Iso3_negative",
                                                 scale(iso3)<  -0.5~"Iso3 peak  Low",
                                                  TRUE~"high_mort_no_peak"
                                                  ))%>% 
  filter(!is.na(get(seriescode)),!is.na(categorie)) -> yo
yo$categorie<-factor(as.factor(yo$categorie),levels=c("Iso3 peak  Low","high_mort_no_peak","Iso3_negative"))
yo2<-yo[,colnames(yo)[colnames(yo)%in% c(paste(seriescode),"categorie")],with=F]
colnames(yo2)[1]<-"value"
series_label<-wdi_explained[wdi_explained$`Series Code` %in% seriescode,"Indicator Name"]

ggstatsplot::ggbetweenstats(
      data = yo2,
      x = categorie,
      y = value ) + 
      # Add labels and title
      labs(x="",
           y = seriescode,
           title = series_label
      )+theme(
        # This is the new default font in the plot
        text = element_text(family = "Roboto", size = 10, color = "black"),
        plot.title = element_text(
          family = "Lobster Two", 
          size = 14,
          face = "bold",
          color = "#2a475e"
        ),
        # # Statistical annotations below the main title
        plot.subtitle = element_text(
          family = "Roboto",
          face = "bold",
          color="#1b2838"
        ),
        plot.title.position = "plot", # slightly different from default
        axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 12)
      )


ggsave("../figures/04_iso3_sig_poverty_boxpplot.png",device = "png")
```
#### Unemployment
```{r}
seriescode<-"SN.ITK.SVFI.ZS"

wdi_data_cons_df %>%  mutate(categorie=case_when(scale(iso3)>1~"Iso3_negative",
                                                 scale(iso3)<  -0.5~"Iso3 peak  Low",
                                                  TRUE~"high_mort_no_peak"
                                                  ))%>% 
  filter(!is.na(get(seriescode)),!is.na(categorie)) -> yo
yo$categorie<-factor(as.factor(yo$categorie),levels=c("Iso3 peak  Low","high_mort_no_peak","Iso3_negative"))
yo2<-yo[,colnames(yo)[colnames(yo)%in% c(paste(seriescode),"categorie")],with=F]
colnames(yo2)[1]<-"value"
series_label<-wdi_explained[wdi_explained$`Series Code` %in% seriescode,"Indicator Name"]

ggstatsplot::ggbetweenstats(
      data = yo2,
      x = categorie,
      y = value ) + 
      # Add labels and title
      labs(x="",
           y = seriescode,
           title = series_label
      )+theme(
        # This is the new default font in the plot
        text = element_text(family = "Roboto", size = 10, color = "black"),
        plot.title = element_text(
          family = "Lobster Two", 
          size = 14,
          face = "bold",
          color = "#2a475e"
        ),
        # # Statistical annotations below the main title
        plot.subtitle = element_text(
          family = "Roboto",
          face = "bold",
          color="#1b2838"
        ),
        plot.title.position = "plot", # slightly different from default
        axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 12)
      )


ggsave("../figures/04_iso3_insecurity.png",device = "png")
```


## iso 4 

```{r,eval=FALSE}
group_values<-wdi_data_cons_df$iso4 %>%  quantile()
thisiso<-paste0("iso4")
var_comb <- expand.grid(var_name = used_var_names,
                          iso_name = thisiso,
                          stringsAsFactors = FALSE)


wdi_data_cons_df %>%  mutate(categorie=case_when(
  iso4<group_values[2]~"0to25",
  iso4<group_values[3]~"25to50",
  iso4<group_values[4]~"50to75",
  iso4<group_values[5]~"75to100",
))-> wdi_data_cons_df

idx_0_25<-wdi_data_cons_df$categorie%in%"0to25"
idx_25_50<-wdi_data_cons_df$categorie%in%"25to50"
idx_50_75<-wdi_data_cons_df$categorie%in%"50to75"
idx_75_100<-wdi_data_cons_df$categorie%in%"75to100"

parallel::clusterEvalQ(cl, suppressWarnings(library(energy)))
parallel::clusterExport(cl, "wdi_data_cons_df", envir = environment())
parallel::clusterExport(cl, "var_comb", envir = environment())
parallel::clusterExport(cl, "idx_0_25", envir = environment())
parallel::clusterExport(cl, "idx_25_50", envir = environment())
parallel::clusterExport(cl, "idx_50_75", envir = environment())
parallel::clusterExport(cl, "idx_75_100", envir = environment())
```


```{r,eval=F}
for(groupA in c("idx_0_25","idx_25_50","idx_50_75","idx_75_100")){
  

variable<-groupA

parallel::clusterExport(cl, "variable", envir = environment())
res <- parLapplyLB(cl, 1:nrow(var_comb), function (i) {
  message("i: ", i, " / ", nrow(var_comb))

  idx <- !is.na(wdi_data_cons_df[[var_comb$var_name[i]]])
  idx<-eval(parse(text=paste0("idx&",variable,sep="")))
  energy::dcor(wdi_data_cons_df[[var_comb$var_name[i]]][idx],
               wdi_data_cons_df[[var_comb$iso_name[i]]][idx])
  
})

res_Tsamples <- parLapplyLB(cl, 1:nrow(var_comb), function (i) {
  message("i: ", i, " / ", nrow(var_comb))
  idx <- !is.na(wdi_data_cons_df[[var_comb$var_name[i]]])
   idx<-eval(parse(text=paste0("idx&",variable,sep="")))
  sum(idx)
})
res_spear <- parLapplyLB(cl, 1:nrow(var_comb), function (i) {
  message("i: ", i, " / ", nrow(var_comb))
  idx <- !is.na(wdi_data_cons_df[[var_comb$var_name[i]]])
   idx<-eval(parse(text=paste0("idx&",variable,sep="")))
  cor(wdi_data_cons_df[[var_comb$var_name[i]]][idx],
               wdi_data_cons_df[[var_comb$iso_name[i]]][idx],method = "pearson")
})

wdi_data_cons_dcor <- matrix(
  NA_real_, length(used_var_names), 1,
  dimnames = list(used_var_names, thisiso)
)
for (i in seq_len(nrow(var_comb))) {
  wdi_data_cons_dcor[var_comb$var_name[i],
                     thisiso] <- res[[i]]
}

wdi_data_cons_dcor2<- matrix(
  NA_real_, length(used_var_names), 1,
  dimnames = list(used_var_names, thisiso)
)
for (i in seq_len(nrow(var_comb))) {
  wdi_data_cons_dcor2[var_comb$var_name[i],
                       thisiso] <- res_spear[[i]]
}
wdi_data_cons_dcor3<- matrix(
  NA_real_, length(used_var_names), 1,
  dimnames = list(used_var_names, thisiso)
)
for (i in seq_len(nrow(var_comb))) {
  wdi_data_cons_dcor3[var_comb$var_name[i],
                       thisiso] <- res_Tsamples[[i]]
}
colnames(wdi_data_cons_dcor2)<-paste("Pearson",colnames(wdi_data_cons_dcor2),sep="_")
colnames(wdi_data_cons_dcor3)<-paste("N",colnames(wdi_data_cons_dcor3),sep="_")

cbind(data.frame(wdi_data_cons_dcor),data.frame(N=wdi_data_cons_dcor3[,1]),wdi_data_cons_dcor2) %>% rownames_to_column(var = "indicators")-> outi
categoriestr<-str_replace(str_remove(groupA,"idx_"),"_","to")





outi$range<- paste0(round(min(wdi_data_cons_df[wdi_data_cons_df$categorie%in%categoriestr,..thisiso]),digits = 3)," to ",
       round(max(wdi_data_cons_df[wdi_data_cons_df$categorie%in%categoriestr,..thisiso]),digits = 3),sep="")
colnames(outi)<-paste(colnames(outi),categoriestr,sep = "_")

eval(parse(text=paste("outi-> WDI_dimred_iso4_",groupA,sep="")))

}

```
### save
```{r,eval=FALSE}
wdi_dimred_iso4<-cbind(WDI_dimred_iso4_idx_0_25,WDI_dimred_iso4_idx_25_50,WDI_dimred_iso4_idx_50_75,WDI_dimred_iso4_idx_75_100)
wdi_dimred_iso4_1<-data.frame("Indicators"=wdi_dimred_iso4$indicators_0to25,wdi_dimred_iso4[,!grepl("indicators",colnames(wdi_dimred_iso4))])



wdi_dimred_iso4_2<-merge(allcorr,wdi_dimred_iso4_1,by="Indicators",all.y=T)
wdi_dimred_iso4_3<-merge(wdi_explained[,c("Series Code","Topic","Indicator Name")],wdi_dimred_iso4_2,by.x="Series Code",by.y="Indicators")
# fwrite(wdi_dimred_iso4_3,"../results/corrTable/04_CorrTable_iso4_quantiles_all.csv")
```

```{r}
iso4_corrtable<-fread("../results/corrTable/04_CorrTable_iso4_quantiles_all.csv")

UPPER25<-quantile(iso4_corrtable$iso4_all_all,na.rm = T)[4]

iso4_corrtable %>% 
            pivot_longer(names_to = "isos",
               c("iso4_all_all","iso4_0to25","iso4_25to50","iso4_50to75","iso4_75to100"),
               values_drop_na = TRUE) %>% 
  ggplot(aes(x=value,fill=isos))+geom_histogram()+geom_vline(xintercept = UPPER25)+ facet_wrap(~isos)
```


#### Textile produciton
```{r}
seriescode<-"TM.VAL.MRCH.R3.ZS"

wdi_data_cons_df %>%  mutate(categorie=case_when(
                                                 scale(iso4)<  -0.6~"ISO_4_high",
                                                  TRUE~"Iso4_normal"
                                                  ))%>% 
  filter(!is.na(get(seriescode)),!is.na(categorie)) -> yo
yo$categorie<-factor(as.factor(yo$categorie),levels=c("ISO_4_high","Iso4_normal"))

yo2<-yo[,colnames(yo)[colnames(yo)%in% c(paste(seriescode),"categorie")],with=F]
colnames(yo2)[1]<-"value"
series_label<-wdi_explained[wdi_explained$`Series Code` %in% seriescode,"Indicator Name"]



ggstatsplot::ggbetweenstats(
      data = yo2,
      x = categorie,
      y = value ) + 
      # Add labels and title
      labs(x="",
           y = seriescode,
           title = series_label
      )+theme(
        # This is the new default font in the plot
        text = element_text(family = "Roboto", size = 10, color = "black"),
        plot.title = element_text(
          family = "Lobster Two", 
          size = 14,
          face = "bold",
          color = "#2a475e"
        ),
        # # Statistical annotations below the main title
        plot.subtitle = element_text(
          family = "Roboto",
          face = "bold",
          color="#1b2838"
        ),
        plot.title.position = "plot", # slightly different from default
        axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 12)
      )


ggsave("../figures/04_iso4_sig_merchandise_boxpplot.png",device = "png",height = 7,width = 9)
```

### population work 


```{r}
seriescode<-"SL.TLF.BASC.MA.ZS"

wdi_data_cons_df %>%  mutate(categorie=case_when(
                                                 scale(iso4)<  -0.6~"ISO_4_high",
                                                  TRUE~"Iso4_normal"
                                                  ))%>% 
  filter(!is.na(get(seriescode)),!is.na(categorie)) -> yo
yo$categorie<-factor(as.factor(yo$categorie),levels=c("ISO_4_high","Iso4_normal"))

yo2<-yo[,colnames(yo)[colnames(yo)%in% c(paste(seriescode),"categorie")],with=F]
colnames(yo2)[1]<-"value"
series_label<-wdi_explained[wdi_explained$`Series Code` %in% seriescode,"Indicator Name"]



ggstatsplot::ggbetweenstats(
      data = yo2,
      x = categorie,
      y = value ) + 
      # Add labels and title
      labs(x="",
           y = seriescode,
           title = series_label
      )+theme(
        # This is the new default font in the plot
        text = element_text(family = "Roboto", size = 10, color = "black"),
        plot.title = element_text(
          family = "Lobster Two", 
          size = 14,
          face = "bold",
          color = "#2a475e"
        ),
        # # Statistical annotations below the main title
        plot.subtitle = element_text(
          family = "Roboto",
          face = "bold",
          color="#1b2838"
        ),
        plot.title.position = "plot", # slightly different from default
        axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 12)
      )


ggsave("../figures/04_iso4_sig_youngworker_boxpplot.png",device = "png",width = 9,height = 7)
```
#### Airpolution
```{r}

seriescode<-"EN.CO2.ETOT.ZS"

wdi_data_cons_df %>%  mutate(categorie=case_when(
                                                 scale(iso4)<  -0.6~"ISO_4_high",
                                                  TRUE~"Iso4_normal"
                                                  ))%>% 
  filter(!is.na(get(seriescode)),!is.na(categorie)) -> yo
yo$categorie<-factor(as.factor(yo$categorie),levels=c("ISO_4_high","Iso4_normal"))

yo2<-yo[,colnames(yo)[colnames(yo)%in% c(paste(seriescode),"categorie")],with=F]
colnames(yo2)[1]<-"value"
series_label<-wdi_explained[wdi_explained$`Series Code` %in% seriescode,"Indicator Name"]



ggstatsplot::ggbetweenstats(
      data = yo2,
      x = categorie,
      y = value ) + 
      # Add labels and title
      labs(x="",
           y = seriescode,
           title = series_label
      )+theme(
        # This is the new default font in the plot
        text = element_text(family = "Roboto", size = 10, color = "black"),
        plot.title = element_text(
          family = "Lobster Two", 
          size = 14,
          face = "bold",
          color = "#2a475e"
        ),
        # # Statistical annotations below the main title
        plot.subtitle = element_text(
          family = "Roboto",
          face = "bold",
          color="#1b2838"
        ),
        plot.title.position = "plot", # slightly different from default
        axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 12)
      )


#ggsave("../figures/04_iso4_sig_youngworker_boxpplot.png",device = "png")
```

## iso 5

```{r,eval=FALSE}
group_values<-wdi_data_cons_df$iso5 %>%  quantile()
thisiso<-paste0("iso5")
var_comb <- expand.grid(var_name = used_var_names,
                          iso_name = thisiso,
                          stringsAsFactors = FALSE)


wdi_data_cons_df %>%  mutate(categorie=case_when(
  iso5<group_values[2]~"0to25",
  iso5<group_values[3]~"25to50",
  iso5<group_values[4]~"50to75",
  iso5<group_values[5]~"75to100",
))-> wdi_data_cons_df

idx_0_25<-wdi_data_cons_df$categorie%in%"0to25"
idx_25_50<-wdi_data_cons_df$categorie%in%"25to50"
idx_50_75<-wdi_data_cons_df$categorie%in%"50to75"
idx_75_100<-wdi_data_cons_df$categorie%in%"75to100"

parallel::clusterEvalQ(cl, suppressWarnings(library(energy)))
parallel::clusterExport(cl, "wdi_data_cons_df", envir = environment())
parallel::clusterExport(cl, "var_comb", envir = environment())
parallel::clusterExport(cl, "idx_0_25", envir = environment())
parallel::clusterExport(cl, "idx_25_50", envir = environment())
parallel::clusterExport(cl, "idx_50_75", envir = environment())
parallel::clusterExport(cl, "idx_75_100", envir = environment())
```


```{r,eval=F}
for(groupA in c("idx_0_25","idx_25_50","idx_50_75","idx_75_100")){
  

variable<-groupA

parallel::clusterExport(cl, "variable", envir = environment())
res <- parLapplyLB(cl, 1:nrow(var_comb), function (i) {
  message("i: ", i, " / ", nrow(var_comb))

  idx <- !is.na(wdi_data_cons_df[[var_comb$var_name[i]]])
  idx<-eval(parse(text=paste0("idx&",variable,sep="")))
  energy::dcor(wdi_data_cons_df[[var_comb$var_name[i]]][idx],
               wdi_data_cons_df[[var_comb$iso_name[i]]][idx])
  
})

res_Tsamples <- parLapplyLB(cl, 1:nrow(var_comb), function (i) {
  message("i: ", i, " / ", nrow(var_comb))
  idx <- !is.na(wdi_data_cons_df[[var_comb$var_name[i]]])
   idx<-eval(parse(text=paste0("idx&",variable,sep="")))
  sum(idx)
})
res_spear <- parLapplyLB(cl, 1:nrow(var_comb), function (i) {
  message("i: ", i, " / ", nrow(var_comb))
  idx <- !is.na(wdi_data_cons_df[[var_comb$var_name[i]]])
   idx<-eval(parse(text=paste0("idx&",variable,sep="")))
  cor(wdi_data_cons_df[[var_comb$var_name[i]]][idx],
               wdi_data_cons_df[[var_comb$iso_name[i]]][idx],method = "pearson")
})

wdi_data_cons_dcor <- matrix(
  NA_real_, length(used_var_names), 1,
  dimnames = list(used_var_names, thisiso)
)
for (i in seq_len(nrow(var_comb))) {
  wdi_data_cons_dcor[var_comb$var_name[i],
                     thisiso] <- res[[i]]
}

wdi_data_cons_dcor2<- matrix(
  NA_real_, length(used_var_names), 1,
  dimnames = list(used_var_names, thisiso)
)
for (i in seq_len(nrow(var_comb))) {
  wdi_data_cons_dcor2[var_comb$var_name[i],
                       thisiso] <- res_spear[[i]]
}
wdi_data_cons_dcor3<- matrix(
  NA_real_, length(used_var_names), 1,
  dimnames = list(used_var_names, thisiso)
)
for (i in seq_len(nrow(var_comb))) {
  wdi_data_cons_dcor3[var_comb$var_name[i],
                       thisiso] <- res_Tsamples[[i]]
}
colnames(wdi_data_cons_dcor2)<-paste("Pearson",colnames(wdi_data_cons_dcor2),sep="_")
colnames(wdi_data_cons_dcor3)<-paste("N",colnames(wdi_data_cons_dcor3),sep="_")

cbind(data.frame(wdi_data_cons_dcor),data.frame(N=wdi_data_cons_dcor3[,1]),wdi_data_cons_dcor2) %>% rownames_to_column(var = "indicators")-> outi
categoriestr<-str_replace(str_remove(groupA,"idx_"),"_","to")





outi$range<- paste0(round(min(wdi_data_cons_df[wdi_data_cons_df$categorie%in%categoriestr,..thisiso]),digits = 3)," to ",
       round(max(wdi_data_cons_df[wdi_data_cons_df$categorie%in%categoriestr,..thisiso]),digits = 3),sep="")
colnames(outi)<-paste(colnames(outi),categoriestr,sep = "_")

eval(parse(text=paste("outi-> WDI_dimred_iso5_",groupA,sep="")))

}

```
### save
```{r,eval=FALSE}
wdi_dimred_iso5<-cbind(WDI_dimred_iso5_idx_0_25,WDI_dimred_iso5_idx_25_50,WDI_dimred_iso5_idx_50_75,WDI_dimred_iso5_idx_75_100)
wdi_dimred_iso5_1<-data.frame("Indicators"=wdi_dimred_iso5$indicators_0to25,wdi_dimred_iso5[,!grepl("indicators",colnames(wdi_dimred_iso5))])



wdi_dimred_iso5_2<-merge(allcorr,wdi_dimred_iso5_1,by="Indicators",all.y=T)
wdi_dimred_iso5_3<-merge(wdi_explained[,c("Series Code","Topic","Indicator Name")],wdi_dimred_iso5_2,by.x="Series Code",by.y="Indicators")
fwrite(wdi_dimred_iso5_3,"../results/corrTable/04_CorrTable_iso5_quantiles_all.csv")
```