---
title: "06_figures_charts"
author: "Kolja Nenoff"
date: "2023-07-08"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
library(rstudioapi)
library(magrittr)
library(tidyverse)
library(data.table)
library(parallel)
library(dimRed)
library(devtools)
library(igraph)
library(energy)
library(ggrepel)
library(ggalt)
library(ggraph)
library(trend)
library(riverplot)
library(RANN)
library(RJSONIO)
library(RSpectra)
library(pcaMethods)
require(broom)
require(GGally)
require(ggpubr)
require(xgboost)
require(caret)
require(pdp)
require(iml)
require(data.table)
require(tidyverse)
require(cowplot)
require(readxl)
```
```{r}
wdi_explained<-fread("../data/WDI_data_230609/WDI_CSV/WDISeries.csv")


dimred2021_knn250_2<-fread("../output/02_230612_Dimred_ExcessMort_annual.csv")
col_pal <- c("#1F968B",  "#7570b3", "#e7298a", "#3CBB75","yellow3", "orange","slategrey")
```
# data
```{r,Excess Mortility WHO Set}
excessCounrty_year<-read_excel("../data/230224_nature_WHO_excess_death/2022-03-25_covid-19_gem/WHO_COVID_Excess_Deaths_EstimatesByCountry.xlsx",sheet = 2)
excessCounrty_covariates<-read_excel("../data/230224_nature_WHO_excess_death/2022-03-25_covid-19_gem/WHO_COVID_Excess_Deaths_EstimatesByCountry.xlsx",sheet = 3)
excessRegion_year<-read_excel("../data/230224_nature_WHO_excess_death/2022-03-25_covid-19_gem/WHO_COVID_Excess_Deaths_EstimatesByRegion.xlsx",sheet = 2)
excessRegion_income<-read_excel("../data/230224_nature_WHO_excess_death/2022-03-25_covid-19_gem/WHO_COVID_Excess_Deaths_EstimatesByIncome.xlsx",sheet = 2)
excessRegion_income %>% group_by(income) %>% 
  summarise(p_value_mort=sum(excess.mean)/sum(expected.mean)*100,excess_comul=sum(excess.mean))
excessCounrty_year_pvalue<-excessCounrty_year %>% group_by(country,iso3,year) %>% 
  summarise(p_value_mort=sum(excess.mean)/sum(expected.mean)*100,excess_comul=sum(excess.mean))
excessCounrty_year %>% filter(year==2020)-> excessCounrty_year_2020
p_value_global_2020=sum(excessCounrty_year_2020$excess.mean,na.rm = T)/sum(excessCounrty_year_2020$expected.mean,na.rm = T)*100
excessCounrty_year %>% filter(year==2021)-> excessCounrty_year_2020
p_value_global_2021=sum(excessCounrty_year_2020$excess.mean,na.rm = T)/sum(excessCounrty_year_2020$expected.mean,na.rm = T)*100
excessCounrty_monthly_pvalue<-excessCounrty_year  %>% 
  mutate(p_value_mort=(excess.mean)/(expected.mean)*100)
```


# Method 1 Compare ISO and PCA
```{r}
## set up path
DATA_DIR <- "../data/"
FIGURE_DIR <- "../figures/01_figures/"
WDI_DIR <- paste0(DATA_DIR, "WDI_data_230609/WDI_CSV")
RELATIVE_INDICATORS_FILE_NAME <- "01_data/relative_indicators_2023_06.R" # nolint
source("../libraries/01a_lib_social_indicators.R")


residual_sample_variances <- readRDS("../data/residual_sample_variances.RDS")
cons_iso_pca<-readRDS("../data/cons_iso_pca.RDS")
occurrence_table       <- cons_iso_pca$occurrence_table
wdi_data_cons          <- cons_iso_pca$wdi_data_cons
wdi_data_cons_pca      <- cons_iso_pca$wdi_data_cons_pca
wdi_data_cons_iso      <- cons_iso_pca$wdi_data_cons_iso
wdi_data_qual_cons_pca <- cons_iso_pca$wdi_data_qual_cons_pca
wdi_data_qual_cons_iso <- cons_iso_pca$wdi_data_qual_cons_iso
knns                   <- cons_iso_pca$knns
knns
cons_iso_pca$wdi_data_cons_iso %>%  head
wdi_data_qual_cons_iso
```


```{r}
pca_res <- residual_sample_variances$pca
iso_res <- residual_sample_variances$iso
median_res_var_pca <- residual_sample_variances$median_pca
median_res_var_iso <- residual_sample_variances$median_iso
hdi_res_var <- residual_sample_variances$hdi

plot_if_no_file(
  get_plot_file("01_230708_pca_iso_res_var.pdf"),
  plot_residual_variances,
  residual_sample_variances = residual_sample_variances,
  wdi_data_qual_cons_iso = wdi_data_qual_cons_iso,
  wdi_data_qual_cons_pca = wdi_data_qual_cons_pca,
  .overwrite = TRUE
)
```
```{r}
residual_sample_variances = residual_sample_variances
wdi_data_qual_cons_pca = wdi_data_qual_cons_pca
wdi_data_qual_cons_iso = wdi_data_qual_cons_iso
  
old_par <- par()
par(mar = c(3.5, 3.75, 0, 0) + 0.1, las = 1, mgp = c(2.5, 0.25, 0))
on.exit(par(old_par))

max_x <- 14

pca_res <- residual_sample_variances$pca
iso_res <- residual_sample_variances$iso
median_res_var_pca <- residual_sample_variances$median_pca
median_res_var_iso <- residual_sample_variances$median_iso
hdi_res_var <- residual_sample_variances$hdi

max_res_var_1 <- 1
png("../figures/06_plots/myplot.png",width = 720,height = 720,units = "px")
plot(pca_res[[1]],
     ylim = c(0, max_res_var_1),
     ## xlim = c(1, max_dim),
     xlim = c(1, max_x),
     ylab = "residual variance",
     xlab = "number of dimensions",
     cex.lab = 2,
     las = 1,
     xaxt = "n",
     yaxt = "n",
     bty = "n",
     type = "n")
points(wdi_data_qual_cons_pca,      type = "b",
       lwd = 4, col = lighten(col_pal[7], 0.8))
points(wdi_data_qual_cons_iso[[2]], type = "b",
       lwd = 4, col = lighten(col_pal[6], 0.8))
text(1, wdi_data_qual_cons_pca[1],      "PCA",
     pos = 4, offset = 0.5, xpd = TRUE, cex = 2)
text(1, wdi_data_qual_cons_iso[[3]][1], "Ensemble Isomap",
     pos = 4, offset = 0.5, xpd = TRUE, cex = 2)
text(max_x + 0.75,         0.1, "10% Residual Variance", cex = 1.8,
     adj = c(1.1, -0.5), xpd = TRUE)
abline(h = 0.1, col = "gray50", lwd = 2)

par(las = 1, mgp = c(1.5, 0.75, 0))#, xpd = TRUE)
axis(side = 1, at = 1:max_x, labels = ifelse(1:14 %% 2 == 1, 1:14, ""),
     col = "gray50", col.axis = "gray50",
     cex.axis = 1.5, lwd = 1, tcl = -0.25, lwd.ticks = 2)
axis(side = 2, at = seq(0, 1, length.out = 6),
     col = "gray50", col.axis = "gray50",
     cex.axis = 1.5, lwd = 1, tcl = -0.25, lwd.ticks = 2)
dev.off
```
# Method 2 Visualisation of  Nonlinearity
```{r}

a<-ggplot(dimred2021_knn250_2) +geom_point(aes(x=iso1,y=SH.STA.MMRT),color=col_pal[1])+ theme_bw() +ylab("Maternal mortality ratio \n(per 100 k live births)")+ xlab("Dimension 1")
b<-ggplot(dimred2021_knn250_2) +geom_point(aes(x=iso1,y=NY.GDP.PCAP.CD),color=col_pal[2])+ theme_bw() +ylab(wdi_explained[wdi_explained$`Series Code`%in% "NY.GDP.PCAP.CD","Indicator Name"])+ xlab("Dimension 1")
library(gridExtra)
grid.arrange(a,b) %>%  ggsave(file="../figures/06_plots/b_methods_nonlinear_pattern.png",device = "png")
```

# Results 1 Shapley values and importance of nonlinear prediction of Xgboost 



```{r}
thisData<-dimred2021_knn250_2[] %>% filter(year%in% 2019)
IsoYear<-"2019"
Label<-"pscore_mean"
candidates<-c("iso1","iso2","iso3","iso4","iso5","iso6","iso7")


custom_control <- trainControl(
  method = "cv",
  number = 30,  # Increase the number of folds
  search = "random",
  allowParallel = TRUE
  
)

param_grid <- expand.grid(
  nrounds = c(50,100),
  eta = c(0.05),
  max_depth = c(2, 4),
  colsample_bytree = c( 1),
  subsample = c(1),
  gamma = c(0),
  min_child_weight = c(0)
)

thisData[thisData$year%in%IsoYear,] %>% select(Label,candidates) %>% drop_na() -> thisData_2
thisData_2 -> thisData_2_input
  ## create and test models

## test different parameters
xgb_caret <- caret::train(x = thisData_2_input %>% dplyr::select(candidates),
                   y = thisData_2_input %>% pull(Label),
                   method = "xgbTree",
                   verbosity = 0,

                   trControl=custom_control,tuneGrid = param_grid,
                   tuneLength = 2
   
                    )

finalmodel <- caret::train(x = thisData_2_input %>% dplyr::select(-Label, ),
                   y = thisData_2_input %>% pull(Label),
                   method = "xgbTree",
                   verbosity = 0,
                   ## CAST CV LLO (Hanna Meyer)
                   trControl=trainControl(method="cv",
                                          number = 30),
                   # early_stopping_rounds = 10,
                   tuneGrid = xgb_caret$bestTune)
finalmodel
```


```{r}

plota<-xgboost::xgb.ggplot.shap.summary(data = as.matrix(thisData_2_input %>% 
  dplyr::select( -Label)),
  model = finalmodel$finalModel) +scale_color_viridis("Feature Value")+ theme_bw() + xlab("Features")+ylab("Shapley value")+theme(legend.position = "left")

subsetimportance<-varImp(finalmodel)$importance %>%
         mutate(var = rownames(varImp(xgb_caret)$importance))%>%
         dplyr::rename(importance = Overall) 


plotb<-ggplot(subsetimportance,aes(x = importance,
    y = reorder(var, importance)),)+
  geom_col(fill="slategrey")+
  ylab("")+theme_classic()+theme(axis.text.y=element_blank(),)

plot_grid(plota, plotb, labels = "A",rel_widths =   c(4/5, 1/5))
#ggsave("../figures/06_plots/ResultsA_nonlinearexmort.png",device = "png")
```
```{r}
importance <- xgb.importance(model = finalmodel$finalModel, data = dtrain)

importance_dt <- data.table::as.data.table(importance)

# Plot the feature importances in a heatmap
xgb.plot.importance(importance_matrix = importance_dt, top_n = 10)
```

# Results 2 WDI or Dimred How predicts better bootstrap

```{r}
result_bootstrap
```

### impute WDI
```{r}

shrinkWDI<-dimred2021_knn250_2[,colnames(dimred2021_knn250_2)%in% c("Country Code","year",wdi_explained$`Series Code`),with=F] %>% filter(year<2020)

aTable<-shrinkWDI[,"Country Code"] %>%  unique
for(i in 3:ncol(shrinkWDI)){
  look<-shrinkWDI[,colnames(shrinkWDI)[c(1,2,i)],with=F] 
  colnames(look)[3]<-"value"
  best<-look[,.(value=last(value),year=last(year)),by=`Country Code`] %>%  filter(!is.na(value))
  okaisch<-look[!(look$`Country Code`%in% best$`Country Code`)&year%in% c(2000:2020),.(value=mean(value,na.rm=T),sd=var(value,na.rm=T)),by=`Country Code`]
  
  thisone<-rbind(best[,c("Country Code","value")],okaisch[,c("Country Code","value")])
  colnames(thisone)[2]<-colnames(shrinkWDI)[c(i)]
  aTable<-left_join(aTable,thisone, by = "Country Code")
}

bTable<-aTable[,-1] %>% sapply( . ,function(x)(as.numeric(x))) %>% data.frame() 

cTable<-sapply(bTable, function(x)x<-if_else(is.nan(x),999,x)) %>% data.frame()

cTable[cTable==999]<-NA
df_filled <- pca(object =  cTable, method = "ppca")
table2019_wdiImputed<-left_join(cbind(aTable[,1],data.frame(df_filled@completeObs)),dimred2021_knn250_2[dimred2021_knn250_2$year%in%2020,c("Country Code","pscore_mean")],by = "Country Code")

```


```{r}

param_grid <- expand.grid(
  nrounds = c(50),
  max_depth = c(1,2,3),           # Maximum depth of a tree
  eta = c(0.05),
  gamma = c(0, 1, 5),               # Minimum loss reduction required to make a further partition on a leaf node (regularization term)
  colsample_bytree = c(1),
  min_child_weight = c(0),
  subsample = c(0.95)
)

custom_control <- trainControl(
  method = "cv",
  number = 60,  # Increase the number of folds
  search = "random",
  allowParallel = TRUE)
  
param_list <- list(objective = "reg:squarederror",  # For regression
                   eta = 0.05,
                   max_depth = c(1,2,3),
                   gamma = c(0, 1, 5),    
                   subsample = 0.95
                   )
```

```{r}
require(doParallel)
data_wdi<-table2019_wdiImputed %>% filter(!is.na(pscore_mean))
trials <- 1000
cl <- makeCluster(30)
registerDoParallel(cl)
# Evaluate the performance of the model using RMSE


system.time(result_bootstrap <- foreach(icount(trials),
                  .combine = bind_rows) %dopar% {
                    require(xgboost)
                    require(tidyverse)
                    require(data.table)
    train_index <- sample(1:nrow(data_wdi), 0.8 * nrow(data_wdi))
    features<-colnames(data_wdi)[!grepl("Country|pscore_mean|[1-9]{2}",colnames(data_wdi))]

    label <- "pscore_mean"
    trainDAT <- data_wdi[train_index, ]
    testDAT <- data_wdi[-train_index, ]
    
    xgb_caret <- caret::train(x = trainDAT %>% dplyr::select(all_of(features)),
                   y = trainDAT %>% pull(label),
                   method = "xgbTree",
                   verbosity = 0,
                   trControl=custom_control,
      tuneGrid = param_grid
                   #,tuneLength = 2
   
                    )
    
    finalmodel <- caret::train(x = trainDAT %>% dplyr::select(all_of(features)),
                   y = trainDAT %>% pull(label),
                   method = "xgbTree",
                   verbosity = 0,
                   ## CAST CV LLO (Hanna Meyer)
                   trControl=trainControl(method="cv",
                                          number = 30),

                   tuneGrid = xgb_caret$bestTune)
    
    importance_matrix<-varImp(finalmodel)$importance %>%
         mutate(var = rownames(varImp(xgb_caret)$importance))%>%
         dplyr::rename(importance = Overall) 
   
    
    thisone<-length(list.files("../test/wdi_bootstrap_rerun/"))+1
    importance_matrix %>% fwrite(paste0("../test/wdi_bootstrap_rerun/run2019toPscoreMean_",thisone,".csv",sep=""))
    
}
)


result_bootstrap_dimred_2019toPScoreMean<-result_bootstrap


#fwrite(result_bootstrap_dimred_2019toPScoreMean,file = "../results/06_wdi_19_to_pscoreMean_bootstrap_1000.csv")
# load("../data/06_wdi_20_bootstrap_1000")


```
### Dimred



```{r}

trials <- 1000
cl <- makeCluster(30)
registerDoParallel(cl)
IsoYear<-"2019"
label<-"pscore_mean"
features<-c("iso1","iso2","iso3","iso4","iso5","iso6","iso7")

dimredData<-dimred2021_knn250_2[] %>% filter(year%in% 2019) %>% select(c(label,features)) %>%  filter(!is.na(iso1),!is.na(pscore_mean))
features<-c("iso1","iso2","iso3","iso4","iso5","iso6","iso7")

label<-"pscore_mean"

system.time(result_bootstrap_dimred <- foreach(icount(trials),
                  .combine = bind_rows) %dopar% {
                    require(xgboost)
                    require(tidyverse)
                    require(data.table)
                    library(shapper)
    
                    
    train_index <- sample(1:nrow(dimredData), 0.8 * nrow(dimredData))
    features<-c("iso1","iso2","iso3","iso4","iso5","iso6","iso7")

    label<-"pscore_mean"
    
    trainDAT <- dimredData[train_index, ]
    testDAT <- dimredData[-train_index, ]
    
    xgb_caret_dimred <- caret::train(x = trainDAT %>% dplyr::select(all_of(features)),
                   y = trainDAT %>% pull(label),
                   method = "xgbTree",
                   verbosity = 0,
                   trControl=custom_control,tuneGrid = param_grid
                   ,tuneLength = 1
   
                    )
    
    
    finalmodel_dimred <- caret::train(x = trainDAT %>% dplyr::select(all_of(features)),
                   y = trainDAT %>% pull(label),
                   method = "xgbTree",
                   verbosity = 0,
                   ## CAST CV LLO (Hanna Meyer)
                   trControl=trainControl(method="cv",
                                          number = 30),
                   # early_stopping_rounds = 10,
                   tuneGrid = xgb_caret_dimred$bestTune)
    
    importance_matrix<-varImp(finalmodel_dimred)$importance %>%
         mutate(var = rownames(varImp(xgb_caret_dimred)$importance))%>%
         dplyr::rename(importance = Overall) 
   

    thisone<-length(list.files("../test/dimred_bootstrap_rerun/"))+1
    importance_matrix %>% fwrite(paste0("../test/dimred_bootstrap_rerun/Dimred_run2019toPscoreMean_",thisone,".csv",sep=""))
    finalmodel_dimred$results
}
)


#fwrite(result_bootstrap_dimred,file = "../results/06_dimred_19_to_pscoreMean_bootstrap_1000.csv")
```
### plot both
```{r}
a<-fread("../results/06_wdi_19_to_pscoreMean_bootstrap_1000.csv")
b<-fread("../results/06_dimred_19_to_pscoreMean_bootstrap_1000.csv")
a$Dataset<-"WDI imputed"
b$Dataset<-"Latent Space"
plotR<-rbind(a,b) %>% ggplot(aes(x=Rsquared,fill=Dataset))+geom_histogram(bins = 100)+theme_classic()+xlim(0,1)+scale_fill_manual(values = col_pal)+theme(legend.position = "sad")+ylab("N")+xlab(expression(paste("R"^2)))
plotRMS<-rbind(a,b) %>% ggplot(aes(x=RMSE,fill=Dataset))+geom_histogram(bins = 100)+theme_classic()+xlim(5,15)+scale_fill_manual(values = col_pal)+theme(legend.position = "sad")+ylab("")

plot_bootstrap<-ggarrange(plotR,plotRMS)
plot_bootstrap_legend<- rbind(a,b) %>% ggplot(aes(x=RMSE,fill=Dataset))+geom_histogram(bins = 100)+scale_fill_manual(values = col_pal)

plot_bootstrap%>%  ggsave(file="../figures/06_plots/c_compare_prediction_bootstrap.png",device = "png",width = 9,height = 4)
plot_bootstrap_legend%>%  ggsave(file="../figures/06_plots/c_compare_prediction_bootstrap_legend.png",device = "png")

```




# Result 3 Monthly excess mort

```{r}
 
excessmonthy <- excessCounrty_monthly_pvalue %>% arrange(country,year,month) %>% rename(`Country Code`=iso3) %>% 
  mutate(date = as.Date(paste(year, month, "01", sep = "-"), format = "%Y-%m-%d"))
# Plot the values
exmortelement<-excessmonthy %>% filter(!is.na(p_value_mort)) %>%  mutate(`P-Score Excess Mortality`=case_when(p_value_mort>50~50,p_value_mort<(-15)~16,TRUE~p_value_mort)) %>% 
 ggplot( . , aes(x = interaction(year, month, lex.order = TRUE), y = (`P-Score Excess Mortality`))) +
  geom_boxplot(aes(fill=as.factor(month))) +theme_classic()+scale_fill_viridis(discrete = T) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x = "", y = "Excess Mortality")+theme(legend.position = "asd")
exmortelement
```

## first round 2019 dimred
```{r}
features<-c("iso1","iso2","iso3","iso4","iso5","iso6","iso7")
dimredData<-dimred2021_knn250_2[] %>% filter(year%in% 2019) %>% select(c(`Country Code`,features)) %>%  filter(!is.na(iso1))
excessmonthy2<-merge(excessmonthy %>% select(date,`Country Code`,p_value_mort),dimredData,by="Country Code")
listofmonth<-excessmonthy$date %>%  unique 
listofmonth<-listofmonth[!is.na(listofmonth)] 
require(xgboost)
require(tidyverse)
require(data.table)
library(shapper)
require("SHAPforxgboost")
cooloutput<-data.frame()

system.time(result_bootstrap_dimred <- for(i in 1:length(listofmonth)){
  



    
    features<-c("iso1","iso2","iso3","iso4","iso5","iso6","iso7")
    thismonth<-listofmonth[i]
    excessmonthy2 %>%  filter(date%in% thismonth) -> runthismonthlysubset
    label<-"pscore_mean"
    
    a<-runthismonthlysubset %>% dplyr::select(all_of(features)) %>% as.matrix()
    xgb_caret_dimred <- caret::train(x = runthismonthlysubset %>% dplyr::select(all_of(features)),
                   y = runthismonthlysubset %>% pull("p_value_mort"),
                   method = "xgbTree",
                   verbosity = 0,trControl=custom_control,tuneGrid = param_grid
                   ,tuneLength = 10
    
                    )
    
    # xgb_caret_dimred
    # mod <- xgboost::xgboost(data = a, 
    #                     label = runthismonthlysubset$p_value_mort, 
    #                     params = param_list, nrounds = 10,
    #                     verbose = FALSE,
    #                     early_stopping_rounds = 8)
    
    finalmodel_dimred <- caret::train(x = runthismonthlysubset %>% dplyr::select(all_of(features)),
                   y = runthismonthlysubset %>% pull(p_value_mort),
                   method = "xgbTree",
                   verbosity = 0,
                   ## CAST CV LLO (Hanna Meyer)
                   trControl=trainControl(method="cv",
                                          number = 60),
                   # early_stopping_rounds = 10,
                   tuneGrid = xgb_caret_dimred$bestTune)
    
    importance_matrix<-varImp(finalmodel_dimred)$importance %>%
         mutate(var = rownames(varImp(xgb_caret_dimred)$importance))%>%
         dplyr::rename(importance = Overall) 
    importance_matrix$name<- thismonth
    importance_matrix$rsquared<-finalmodel_dimred$results$Rsquared
    
    # shap_values <- shap.values(xgb_model = mod, X_train = a)
    # # The ranked features by mean |SHAP|
    # shaploutput<-shap_values$mean_shap_score 
    # 
    # orderShap<-match(rownames(importance_matrix), names(shaploutput))
    # 
    # importance_matrix$shapmean<-shaploutput[orderShap]
    cooloutput<-rbind(cooloutput,importance_matrix)
    
})

```
```{r}

rsquared_values_plot<-
  cooloutput %>% select(name,rsquared) %>%  unique %>% ggplot( . , aes(x=name, y=rsquared)) +
  geom_segment( aes(x=name, xend=name, y=0, yend=rsquared), color="grey") +
  geom_point( color=col_pal[5], size=4) +
  theme_light() + geom_line()+
  theme(
    panel.grid.major.x = element_blank(),
    panel.border = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  xlab("") +
  ylab("Value of Y") +ylim(0.2,1)
# heatmap
```


```{r}


start_date<- as.Date("2020-1-1")
end_date<- as.Date("2021-12-1")
# Heatmap 
importanceplot<-cooloutput %>% filter(!is.na(var))%>% 
  mutate(var = factor(var, levels=rev(features))) %>% 
  ggplot( . , aes(name, var, fill= importance)) + 
  geom_tile()+scale_fill_viridis()+theme_classic()+xlab("")+xlim(1,1)+
  scale_x_date(limits = c(start_date, end_date))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+theme(legend.position = "sad")

plot_grid(rsquared_values_plot,exmortelement, importanceplot,ncol = 1,rel_heights =     c(1.5/5, 2/5,3/5))

```



## second round 2020 and 2021

### looks horrible dont do it again
```{r}

features<-c("iso1","iso2","iso3","iso4","iso5","iso6","iso7")
dimredData2020<-dimred2021_knn250_2[] %>% filter(year%in% 2020) %>% 
  mutate(identifier=paste0(`Country Code`,"_",year)) %>% 
  select(c(identifier,features)) %>%  filter(!is.na(iso1)) 
dimredData2021<-dimred2021_knn250_2[] %>% filter(year%in% 2021) %>% 
  mutate(identifier=paste0(`Country Code`,"_",year)) %>% 
  select(c(identifier,features)) %>%  filter(!is.na(iso1)) 
dimredDataboth<-rbind(dimredData2020,dimredData2021)
excessmonthy<-excessmonthy %>% 
  mutate(identifier=paste0(`Country Code`,"_",year))
excessmonthy2<-merge(excessmonthy,dimredDataboth,by="identifier")
listofmonth<-excessmonthy$date %>%  unique 
listofmonth<-listofmonth[!is.na(listofmonth)] 
require(xgboost)
require(tidyverse)
require(data.table)
library(shapper)
require("SHAPforxgboost")
cooloutput_2020_2021<-data.frame()

system.time(result_bootstrap_dimred <- for(i in 1:length(listofmonth)){
  



    
    features<-c("iso1","iso2","iso3","iso4","iso5","iso6","iso7")
    thismonth<-listofmonth[i]
    excessmonthy2 %>%  filter(date%in% thismonth) -> runthismonthlysubset
    label<-"pscore_mean"
    
    a<-runthismonthlysubset %>% dplyr::select(all_of(features)) %>% as.matrix()
    xgb_caret_dimred <- caret::train(x = runthismonthlysubset %>% dplyr::select(all_of(features)),
                   y = runthismonthlysubset %>% pull("p_value_mort"),
                   method = "xgbTree",
                   verbosity = 0,trControl=custom_control,tuneGrid = param_grid
                   ,tuneLength = 10
    
                    )
    
    # xgb_caret_dimred
    # mod <- xgboost::xgboost(data = a, 
    #                     label = runthismonthlysubset$p_value_mort, 
    #                     params = param_list, nrounds = 10,
    #                     verbose = FALSE,
    #                     early_stopping_rounds = 8)
    
    finalmodel_dimred <- caret::train(x = runthismonthlysubset %>% dplyr::select(all_of(features)),
                   y = runthismonthlysubset %>% pull(p_value_mort),
                   method = "xgbTree",
                   verbosity = 0,
                   ## CAST CV LLO (Hanna Meyer)
                   trControl=trainControl(method="cv",
                                          number = 30),
                   # early_stopping_rounds = 10,
                   tuneGrid = xgb_caret_dimred$bestTune)
    
    importance_matrix<-varImp(finalmodel_dimred)$importance %>%
         mutate(var = rownames(varImp(xgb_caret_dimred)$importance))%>%
         dplyr::rename(importance = Overall) 
    importance_matrix$name<- thismonth
    importance_matrix$rsquared<-finalmodel_dimred$results$Rsquared
    
    # shap_values <- shap.values(xgb_model = mod, X_train = a)
    # # The ranked features by mean |SHAP|
    # shaploutput<-shap_values$mean_shap_score 
    # 
    # orderShap<-match(rownames(importance_matrix), names(shaploutput))
    # 
    # importance_matrix$shapmean<-shaploutput[orderShap]
    cooloutput_2020_2021<-rbind(cooloutput_2020_2021,importance_matrix)
    
})
```
```{r}

rsquared_values_plot<-
  cooloutput_2020_2021 %>% select(name,rsquared) %>%  unique %>% ggplot( . , aes(x=name, y=rsquared)) +
  geom_segment( aes(x=name, xend=name, y=0, yend=rsquared), color="grey") +
  geom_point( color=col_pal[5], size=4) +
  theme_light() + geom_line()+
  theme(
    panel.grid.major.x = element_blank(),
    panel.border = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  xlab("") +
  ylab("Value of Y") +ylim(0.2,1)
# heatmap
```


```{r}


start_date<- as.Date("2020-1-1")
end_date<- as.Date("2021-12-1")
# Heatmap 
importanceplot<-cooloutput_2020_2021 %>% filter(!is.na(var))%>% 
  mutate(var = factor(var, levels=rev(features))) %>% 
  ggplot( . , aes(name, var, fill= importance)) + 
  geom_tile()+scale_fill_viridis()+theme_classic()+xlab("")+xlim(1,1)+
  scale_x_date(limits = c(start_date, end_date))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+theme(legend.position = "sad")

plot_grid(rsquared_values_plot,exmortelement, importanceplot,ncol = 1,rel_heights =     c(1.5/5, 2/5,3/5))

```
## third round 2019 dimred on clustes
### the clusters
```{r}
clusterbyExcessandIsos<-dimred2021_knn250_2 %>% filter(year%in% 2019,`Country Code`!="PER") %>%  select(`Country Code`,pscore_mean,features) %>%  drop_na()
kmeans_result <- kmeans(clusterbyExcessandIsos[,-1], centers = 5)
clusterbyExcessandIsos$Cluster<-kmeans_result$cluster
```

```{r}
theclusteredFrame<-data.frame()
for(i in unique(clusterbyExcessandIsos$Cluster))
  
  
  

var_caountry<-clusterbyExcessandIsos[clusterbyExcessandIsos$Cluster %in% i,] %>% select(`Country Code`) %>%  unlist()
features<-c("iso1","iso2","iso3","iso4","iso5","iso6","iso7")
dimredData<-dimred2021_knn250_2[] %>% filter(year%in% 2019) %>% filter(`Country Code`%in% var_caountry) %>% select(c(`Country Code`,features)) %>%  filter(!is.na(iso1))
excessmonthy2<-merge(excessmonthy %>% select(date,`Country Code`,p_value_mort),dimredData,by="Country Code")

listofmonth<-excessmonthy$date %>%  unique 
listofmonth<-listofmonth[!is.na(listofmonth)] 
require(xgboost)
require(tidyverse)
require(data.table)
library(shapper)
require("SHAPforxgboost")
cooloutput<-data.frame()

system.time(result_bootstrap_dimred <- for(i in 1:length(listofmonth)){
  

    
    features<-c("iso1","iso2","iso3","iso4","iso5","iso6","iso7")
    thismonth<-listofmonth[i]
    excessmonthy2 %>%  filter(date%in% thismonth) -> runthismonthlysubset
    label<-"pscore_mean"
    
    a<-runthismonthlysubset %>% dplyr::select(all_of(features)) %>% as.matrix()
    xgb_caret_dimred <- caret::train(x = runthismonthlysubset %>% dplyr::select(all_of(features)),
                   y = runthismonthlysubset %>% pull("p_value_mort"),
                   method = "xgbTree",
                   verbosity = 0,trControl=custom_control,tuneGrid = param_grid
                   ,tuneLength = 10
    
                    )
    
    # xgb_caret_dimred
    # mod <- xgboost::xgboost(data = a, 
    #                     label = runthismonthlysubset$p_value_mort, 
    #                     params = param_list, nrounds = 10,
    #                     verbose = FALSE,
    #                     early_stopping_rounds = 8)
    
    finalmodel_dimred <- caret::train(x = runthismonthlysubset %>% dplyr::select(all_of(features)),
                   y = runthismonthlysubset %>% pull(p_value_mort),
                   method = "xgbTree",
                   verbosity = 0,
                   ## CAST CV LLO (Hanna Meyer)
                   trControl=trainControl(method="cv",
                                          number = 60),
                   # early_stopping_rounds = 10,
                   tuneGrid = xgb_caret_dimred$bestTune)
    
    importance_matrix<-varImp(finalmodel_dimred)$importance %>%
         mutate(var = rownames(varImp(xgb_caret_dimred)$importance))%>%
         dplyr::rename(importance = Overall) 
    importance_matrix$name<- thismonth
    importance_matrix$rsquared<-finalmodel_dimred$results$Rsquared
    
    # shap_values <- shap.values(xgb_model = mod, X_train = a)
    # # The ranked features by mean |SHAP|
    # shaploutput<-shap_values$mean_shap_score 
    # 
    # orderShap<-match(rownames(importance_matrix), names(shaploutput))
    # 
    # importance_matrix$shapmean<-shaploutput[orderShap]
    cooloutput<-rbind(cooloutput,importance_matrix)
    
})
cooloutput$cluster<-i
theclusteredFrame<-bind_rows(theclusteredFrame,cooloutput)

}
```
```{r}

rsquared_values_plot<-
  cooloutput %>% select(name,rsquared) %>%  unique %>% ggplot( . , aes(x=name, y=rsquared)) +
  geom_segment( aes(x=name, xend=name, y=0, yend=rsquared), color="grey") +
  geom_point( color=col_pal[5], size=4) +
  theme_light() + geom_line()+
  theme(
    panel.grid.major.x = element_blank(),
    panel.border = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  xlab("") +
  ylab("Value of Y") +ylim(0.2,1)
# heatmap
```


```{r}


start_date<- as.Date("2020-1-1")
end_date<- as.Date("2021-12-1")
# Heatmap 
importanceplot<-cooloutput %>% filter(!is.na(var))%>% 
  mutate(var = factor(var, levels=rev(features))) %>% 
  ggplot( . , aes(name, var, fill= importance)) + 
  geom_tile()+scale_fill_viridis()+theme_classic()+xlab("")+xlim(1,1)+
  scale_x_date(limits = c(start_date, end_date))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+theme(legend.position = "sad")

plot_grid(rsquared_values_plot,exmortelement, importanceplot,ncol = 1,rel_heights =     c(1.5/5, 2/5,3/5))

```


```{r}
excessmonthy %>% filter(date>"2020-02-01")
excessmonthy %>%  filter(`Country Code`%in% "PER") %>% ggplot(aes(x=date,y=p_value_mort))+ geom_point()+ geom_line()
excessmonthy2<-merge(excessmonthy %>% select(date,`Country Code`,p_value_mort),dimred2021_knn250_2,by="Country Code")
```

