---
title: "07_figures_charts_Part2"
author: "Kolja Nenoff"
date: "2023-08-24"
output: html_document
---
# Results 5 circular plot Signficant correlated WDI features in a 
`
## PLOT 1 all data
```{r}
DATA_DIR<-"../data/"
RELATIVE_INDICATORS_FILE_NAME <- "../data/01_data/relative_indicators_2023_06.R"

order_indidactors<-iso1_onlyDimred$`Series Code` 
dcor_var_iso_cons<-matrix(nrow =length(order_indidactors),ncol=7)
rownames(dcor_var_iso_cons)<-order_indidactors
dcor_var_iso_cons[,1]<-iso1_onlyDimred[iso1_onlyDimred$`Series Code`%in% order_indidactors,"iso1_all"] %>%  unlist
dcor_var_iso_cons[,2]<-iso2_onlyDimred[iso2_onlyDimred$`Series Code`%in% order_indidactors,"iso2_all"] %>%  unlist
dcor_var_iso_cons[,3]<-iso3_onlyDimred[iso3_onlyDimred$`Series Code`%in% order_indidactors,"iso3_all"] %>%  unlist
dcor_var_iso_cons[,4]<-iso4_onlyDimred[iso4_onlyDimred$`Series Code`%in% order_indidactors,"iso4_all"] %>%  unlist
dcor_var_iso_cons[,5]<-iso5_onlyDimred[iso5_onlyDimred$`Series Code`%in% order_indidactors,"iso5_all"] %>%  unlist
dcor_var_iso_cons[,6]<-iso6_onlyDimred[iso6_onlyDimred$`Series Code`%in% order_indidactors,"iso6_all"] %>%  unlist
dcor_var_iso_cons[,7]<-iso7_onlyDimred[iso7_onlyDimred$`Series Code`%in% order_indidactors,"iso7_all"] %>%  unlist


used_var_names_max_dcor <- c(
  dcor_var_iso_cons[, 1] %>% order(decreasing = TRUE) %>% { .[1:15] } %>% { names(dcor_var_iso_cons[., 1]) },
  dcor_var_iso_cons[, 2] %>% order(decreasing = TRUE) %>% { .[1:15] } %>% { names(dcor_var_iso_cons[., 2]) },
  dcor_var_iso_cons[, 3] %>% order(decreasing = TRUE) %>% { .[1:30] } %>% { names(dcor_var_iso_cons[., 3]) },
  dcor_var_iso_cons[, 4] %>% order(decreasing = TRUE) %>% { .[1:30] } %>% { names(dcor_var_iso_cons[., 4]) },
  dcor_var_iso_cons[, 5] %>% order(decreasing = TRUE) %>% { .[1:05] } %>% { names(dcor_var_iso_cons[., 5]) },
  dcor_var_iso_cons[, 6] %>% order(decreasing = TRUE) %>% { .[1:05] } %>% { names(dcor_var_iso_cons[., 6]) },
  dcor_var_iso_cons[, 7] %>% order(decreasing = TRUE) %>% { .[1:10]  } %>% {names(dcor_var_iso_cons[., 7]) }
) %>% unique %>% sort
subsetwdi<-wdi_explained[wdi_explained$`Series Code`%in% used_var_names_max_dcor,]

```

#### input
```{r}
source("../libraries/01a_lib_social_indicators.R")
DATA_DIR<-"../data/"
RELATIVE_INDICATORS_FILE_NAME <- "../data/01_data/relative_indicators_2023_06.R"
var_names         = used_var_names_max_dcor
dcor_var_iso_cons = dcor_var_iso_cons
load("../data/wdi_series")
local_wdi_series  = wdi_series[,]



```

#### setup
```{r}
subsetwdi<-wdi_explained[wdi_explained$`Series Code`%in% used_var_names_max_dcor,]

var_names         = used_var_names_max_dcor
dcor_var_iso_cons = dcor_var_iso_cons
load("../data/wdi_series")
local_wdi_series  = wdi_series[,]
expand_multiplier = 1 / 10
text_size         = 5
label_modifier    = manual_label_modifier
col_pal <- c("#1F968B", "#3CBB75", "#7570b3", "#e7298a", "yellow3", "orange","slategrey")
```

### plot function
```{r}

#label_modifier    = short_name_label_modifier,
## .overwrite = TRUE, # nolint: commented_code_linter.
.width  = 15
.height = 15

component_common <- "Dimension"
component_names <- paste(component_common, 1:7)
xy_expand <- function(xy, expand) xy * (1 + expand * expand_multiplier)
# Does not work
local_wdi_series<-local_wdi_series[grepl(":",local_wdi_series$Topic),]

local_wdi_series <- wdi_series_split_topic(local_wdi_series)

wdi_series_used_important <- local_wdi_series[`Indicator Code` %in% var_names]
var_names<-var_names[var_names%in% wdi_series_used_important$`Indicator Code`]
require(igraph)
stopifnot(length(var_names) == nrow(wdi_series_used_important))
dcor_var_iso_cons_subset<-dcor_var_iso_cons[rownames(dcor_var_iso_cons)%in% var_names,]
dcor_var_iso_cons<-dcor_var_iso_cons[rownames(dcor_var_iso_cons)%in% var_names,]



local_wdi_series<-local_wdi_series[local_wdi_series$`Indicator Code`%in% var_names,]
require(tidygraph)
require(ggraph)

dcor_subset<-dcor_var_iso_cons_subset
## local_wdi_series$`Indicator Code`
edges_level_01 <- local_wdi_series %>% select(`Indicator Code`, Topic1) %>%
  unique %>% rename( from = Topic1, to = `Indicator Code`)
edges_level_12 <- local_wdi_series %>% select(Topic1, Topic2)           %>%
  unique %>% rename( from = Topic2, to = Topic1          )
edges_level_23 <- local_wdi_series %>% select(Topic2, Topic3)           %>%
  unique %>% rename( from = Topic3, to = Topic2          )
edges_level_34 <- local_wdi_series %>% select(Topic3, Topic4)           %>%
  unique %>% rename( from = Topic4, to = Topic3          )
edges_level_45 <- local_wdi_series %>% select(Topic4, Topic5)           %>%
  unique %>% rename( from = Topic5, to = Topic4          )
## create the actual igraph object

rbind(edges_level_01, edges_level_12, edges_level_23, edges_level_34,
      edges_level_45) %>%
unique %>%
filter(to != from) ->listi
hier_graph <-listi %>% data.frame() %>%  { .[rev(seq_len(ncol(.)))] } %>% graph_from_data_frame
  ## graph_from_data_frame does not care about column names,
  ## only about positions
  # %>%
  #graph_from_data_frame

## additional data:

graph_idxs <- match(var_names, V(hier_graph)$name)

hier_graph <- set_vertex_attr(hier_graph, "which_max_iso", graph_idxs,
                              dcor_subset %>% apply(1, which.max))
hier_graph <- set_vertex_attr(hier_graph, "dcor_max", graph_idxs,
                              dcor_subset %>% apply(1, max))
hier_graph <- set_vertex_attr(hier_graph, "dcor1", graph_idxs,
                              dcor_subset[, 1])
hier_graph <- set_vertex_attr(hier_graph, "dcor2", graph_idxs,
                              dcor_subset[, 2])
hier_graph <- set_vertex_attr(hier_graph, "dcor3", graph_idxs,
                              dcor_subset[, 3])
hier_graph <- set_vertex_attr(hier_graph, "dcor4", graph_idxs,
                              dcor_subset[, 4])
hier_graph <- set_vertex_attr(hier_graph, "dcor5", graph_idxs,
                              dcor_subset[, 5])
hier_graph <- set_vertex_attr(hier_graph, "dcor6", graph_idxs,
                              dcor_subset[, 6])
hier_graph <- set_vertex_attr(hier_graph, "dcor7", graph_idxs,
                              dcor_subset[, 7])
# 

hier_graph <-
  set_vertex_attr(hier_graph, "Indicator.Code", graph_idxs,
                  local_wdi_series$`Indicator Code`[
                               match(var_names, local_wdi_series$`Indicator Code`)])



# thisPeople<-fread("../data/indicator_compressed_names.csv")
# wdi_names_prepared<-wdi_series_used_important[,c("Indicator Code","Indicator Name")]
# 
# 
# merge(wdi_names_prepared,thisPeople,by="Indicator Code",all.x = T)->wdi_names_prepared2

# wdi_names_prepared2 %>% select(`Indicator Code`,`Indicator Name`,`Compressed Name`)%>%  filter(is.na(`Compressed Name`)) %>% fwrite("../output/preparethenameschatgpt.csv")
# 
# wdi_series_used[,c("Indicator Code","Compressed Name")] #%>% fwrite("../data/indicator_compressed_names.csv")
# # # chatgpt shortened the names
# wdi_series_used
# wdi_series_used2<-fread("../data/indicator_compressed_names.csv")
# shortnames<-fread("../data/05_abbrevationmaker.csv",sep=";")
# rbind(wdi_series_used2[!(wdi_series_used2$`Indicator Code`%in% shortnames$`Indicator Code`)],shortnames) %>%  fwrite("../output/05_indicator_compressed_names.csv")

wdi_series_used<-fread("../data/indicator_compressed_names.csv")

wdi_var_hierarchy_graph <-hier_graph
axis_cons <- lapply(1:7, inner_connections_by_axis,
                       dcor_var_iso_cons       = dcor_var_iso_cons,
                       wdi_var_hierarchy_graph = wdi_var_hierarchy_graph,
                       var_names               = var_names)




expand_multiplier = 1/2
text_size         = 3
label_modifier    = manual_label_modifier
#label_modifier    = short_name_label_modifier
## .overwrite = TRUE, # nolint: commented_code_linter.
.width  = 50
.height = 50


wdi_var_hierarchy_graph %>%
ggraph(layout = "dendrogram", circular = TRUE) +
geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor1),
                    y = xy_expand(y, dcor1)),
                colour = col_pal[1]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor1),
                 y = xy_expand(y, dcor1)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[1]) +
geom_node_point(aes(filter = leaf,
                   x = xy_expand(x, dcor2),
                    y = xy_expand(y, dcor2)),
                colour = col_pal[2]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor2),
                 y = xy_expand(y, dcor2)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[2]) +
geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor3),
                    y = xy_expand(y, dcor3)),
                colour = col_pal[3]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor3),
                 y = xy_expand(y, dcor3)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[3]) +
geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor4),
                    y = xy_expand(y, dcor4)),
                colour = col_pal[4]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor4),
                 y = xy_expand(y, dcor4)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[4]) +
geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor5),
                    y = xy_expand(y, dcor5)),
                colour = col_pal[5]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor5),
                 y = xy_expand(y, dcor5)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[5]) +
  geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor5),
                    y = xy_expand(y, dcor5)),
                colour = col_pal[5]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor5),
                 y = xy_expand(y, dcor5)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[5]) +
  geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor5),
                    y = xy_expand(y, dcor5)),
                colour = col_pal[6]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor5),
                 y = xy_expand(y, dcor5)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[6]) +
      geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor5),
                    y = xy_expand(y, dcor5)),
                colour = col_pal[7]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor5),
                 y = xy_expand(y, dcor5)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[7]) +
geom_node_point(aes(filter = leaf, color = factor(which_max_iso)),
                size = 0)+
geom_node_text(aes(filter = leaf,
                   label  = label_modifier(as.character(Indicator.Code)),
                   color  = factor(paste(component_common, which_max_iso)),
                   x      = x * (1 + 1 * expand_multiplier),
                   y      = y * (1 + 1 * expand_multiplier),
                   hjust  = if_else(x > 0, 0, 1),
                   angle  = atan(y / x) / 2 / pi * 360),
               show.legend = FALSE,
               size        = text_size) +
geom_conn_bundle(data = get_con(axis_cons[[1]][1, ], axis_cons[[1]][2, ]),
                 edge_colour = col_pal[1], edge_alpha = 0.3) +
geom_conn_bundle(data = get_con(axis_cons[[2]][1, ], axis_cons[[2]][2, ]),
                 edge_colour = col_pal[2], edge_alpha = 0.3) +
geom_conn_bundle(data = get_con(axis_cons[[3]][1, ], axis_cons[[3]][2, ]),
                 edge_colour = col_pal[3], edge_alpha = 0.3) +
(
  if (ncol(axis_cons[[4]]) > 0)
    geom_conn_bundle(data = get_con(axis_cons[[4]][1, ],
                                    axis_cons[[4]][2, ]),
                     edge_colour = col_pal[4], edge_alpha = 0.3)
  else
    list()
) +
geom_conn_bundle(data = get_con(axis_cons[[5]][1, ], axis_cons[[5]][2, ]),
                 edge_colour = col_pal[5], edge_alpha = 0.3) +
geom_edge_diagonal() +
# geom_node_label(aes(filter = !leaf,
#                     fontface = if_else(name %in% level1, "bold", "plain"),
#                     label = name, angle = atan(y / x) / 2 / pi * 360),
#                 size = 2) +
scale_color_manual(limits = component_names,
                   labels = component_names,
                   values = col_pal) +
guides(colour = guide_legend(title = component_common,
                             override.aes = list(size = 5))) +
theme_void() +
theme(legend.title = element_blank(),
      legend.position = c(0.9, 0.1),
      legend.justification = c(0.9, 0.1)) +
expand_limits(x = c(-2.7, 2.7), y = c(-2.7, 2.7))
# 
ggsave("../figures/07_230808_circle_all.pdf",device = "pdf",height = 12,width = 12)

```
#### Description
```{r}
subsetwdi$Topic %>%  table %>%  data.frame() %>%  arrange(desc(Freq))
```

## PLOT 2 SUBSET Significant shapleys


### input
#### data

```{r}
DATA_DIR<-"../data/"
RELATIVE_INDICATORS_FILE_NAME <- "../data/01_data/relative_indicators_2023_06.R"

order_indidactors<-iso1_onlyDimred$`Series Code` 
dcor_var_iso_cons<-matrix(nrow =length(order_indidactors),ncol=7)
rownames(dcor_var_iso_cons)<-order_indidactors

iso1_onlyDimred %>% filter(N>=mean(N))->iso1_onlyDimred_2
iso2_onlyDimred %>% filter(N_shapley>=mean(N_shapley))->iso2_onlyDimred_2
iso3_onlyDimred %>% filter(N_shapley>=mean(N_shapley))->iso3_onlyDimred_2
iso4_onlyDimred %>% filter(N_shapley>=mean(N_shapley))->iso4_onlyDimred_2
iso5_onlyDimred %>% filter(N_shapley>=mean(N_shapley))->iso5_onlyDimred_2
iso6_onlyDimred %>% filter(N_shapley>=mean(N_shapley))->iso6_onlyDimred_2
iso7_onlyDimred %>% filter(N_shapley>=mean(N_shapley))->iso7_onlyDimred_2

theseindicators<-unique(c(iso1_onlyDimred_2$`Series Code`,iso2_onlyDimred_2$`Series Code`,iso3_onlyDimred_2$`Series Code`,
                          iso4_onlyDimred_2$`Series Code`,iso5_onlyDimred_2$`Series Code`,iso6_onlyDimred_2$`Series Code`,
                          iso7_onlyDimred_2$`Series Code`))

dcor_var_iso_cons[,1]<-iso1_onlyDimred[iso1_onlyDimred$`Series Code`%in% order_indidactors,"iso1"] %>%  unlist
dcor_var_iso_cons[,2]<-iso2_onlyDimred[iso2_onlyDimred$`Series Code`%in% order_indidactors,"iso2_75to100_75to100"] %>%  unlist
dcor_var_iso_cons[,3]<-iso3_onlyDimred[iso3_onlyDimred$`Series Code`%in% order_indidactors,"iso3_shapley"] %>%  unlist
dcor_var_iso_cons[,4]<-iso4_onlyDimred[iso4_onlyDimred$`Series Code`%in% order_indidactors,"iso4_0to25_0to25"] %>%  unlist
dcor_var_iso_cons[,5]<-iso5_onlyDimred[iso5_onlyDimred$`Series Code`%in% order_indidactors,"iso5_shapley"] %>%  unlist
dcor_var_iso_cons[,6]<-iso6_onlyDimred[iso6_onlyDimred$`Series Code`%in% order_indidactors,"iso6_shapley"] %>%  unlist
dcor_var_iso_cons[,7]<-iso7_onlyDimred[iso7_onlyDimred$`Series Code`%in% order_indidactors,"iso7_shapley"] %>%  unlist
dcor_var_iso_cons<-dcor_var_iso_cons[rownames(dcor_var_iso_cons) %in%theseindicators,]
dcor_var_iso_cons %>% 
```

#### setup
```{r}


used_var_names_max_dcor <- c(
  dcor_var_iso_cons[, 1] %>% order(decreasing = TRUE) %>% { .[1:15] } %>% { names(dcor_var_iso_cons[., 1]) },
  dcor_var_iso_cons[, 2] %>% order(decreasing = TRUE) %>% { .[1:20] } %>% { names(dcor_var_iso_cons[., 2]) },
  dcor_var_iso_cons[, 3] %>% order(decreasing = TRUE) %>% { .[1:30] } %>% { names(dcor_var_iso_cons[., 3]) },
  dcor_var_iso_cons[, 4] %>% order(decreasing = TRUE) %>% { .[1:30] } %>% { names(dcor_var_iso_cons[., 4]) },
  dcor_var_iso_cons[, 5] %>% order(decreasing = TRUE) %>% { .[1:05] } %>% { names(dcor_var_iso_cons[., 5]) },
  dcor_var_iso_cons[, 6] %>% order(decreasing = TRUE) %>% { .[1:05] } %>% { names(dcor_var_iso_cons[., 6]) },
  dcor_var_iso_cons[, 7] %>% order(decreasing = TRUE) %>% { .[1:05]  } %>% {names(dcor_var_iso_cons[., 7]) }
) %>% unique %>% sort

subsetwdi<-wdi_explained[wdi_explained$`Series Code`%in% used_var_names_max_dcor,]

var_names         = used_var_names_max_dcor
dcor_var_iso_cons = dcor_var_iso_cons
load("../data/wdi_series")
local_wdi_series  = wdi_series[,]
expand_multiplier = 1 / 10
text_size         = 5
label_modifier    = manual_label_modifier
col_pal <- c("#1F968B", "#3CBB75", "#7570b3", "#e7298a", "yellow3", "orange","slategrey")
```

#### modifiy the compressed names
```{r}
compressedNames <- fread("../data/indicator_compressed_names.csv")

missingIndicators <- wdi_explained %>%
  filter(`Series Code` %in% used_var_names_max_dcor) %>%
  select(`Series Code`,`Indicator Name`) %>%
  rename(`Indicator Code` = `Series Code`) %>%
  merge(compressedNames[,c("Indicator Code","Compressed Name")], by = "Indicator Code", all.x = TRUE) %>%
  filter(is.na(`Compressed Name`))

missingIndicators

# 
#missingIndicators %>% fwrite("../data/05_abbrevationmaker.csv",sep=";")

#rbind(compressedNames,addthis) %>% fwrite("../data/indicator_compressed_names.csv")
```



### plot function
```{r}

#label_modifier    = short_name_label_modifier,
## .overwrite = TRUE, # nolint: commented_code_linter.
.width  = 15
.height = 15

component_common <- "Dimension"
component_names <- paste(component_common, 1:7)
xy_expand <- function(xy, expand) xy * (1 + expand * expand_multiplier)
# Does not work
local_wdi_series<-local_wdi_series[grepl(":",local_wdi_series$Topic),]

local_wdi_series <- wdi_series_split_topic(local_wdi_series)

wdi_series_used_important <- local_wdi_series[`Indicator Code` %in% var_names]
var_names<-var_names[var_names%in% wdi_series_used_important$`Indicator Code`]
require(igraph)
stopifnot(length(var_names) == nrow(wdi_series_used_important))
dcor_var_iso_cons_subset<-dcor_var_iso_cons[rownames(dcor_var_iso_cons)%in% var_names,]
dcor_var_iso_cons<-dcor_var_iso_cons[rownames(dcor_var_iso_cons)%in% var_names,]



local_wdi_series<-local_wdi_series[local_wdi_series$`Indicator Code`%in% var_names,]
require(tidygraph)
require(ggraph)

dcor_subset<-dcor_var_iso_cons_subset
## local_wdi_series$`Indicator Code`
edges_level_01 <- local_wdi_series %>% select(`Indicator Code`, Topic1) %>%
  unique %>% rename( from = Topic1, to = `Indicator Code`)
edges_level_12 <- local_wdi_series %>% select(Topic1, Topic2)           %>%
  unique %>% rename( from = Topic2, to = Topic1          )
edges_level_23 <- local_wdi_series %>% select(Topic2, Topic3)           %>%
  unique %>% rename( from = Topic3, to = Topic2          )
edges_level_34 <- local_wdi_series %>% select(Topic3, Topic4)           %>%
  unique %>% rename( from = Topic4, to = Topic3          )
edges_level_45 <- local_wdi_series %>% select(Topic4, Topic5)           %>%
  unique %>% rename( from = Topic5, to = Topic4          )
## create the actual igraph object

rbind(edges_level_01, edges_level_12, edges_level_23, edges_level_34,
      edges_level_45) %>%
unique %>%
filter(to != from) ->listi
hier_graph <-listi %>% data.frame() %>%  { .[rev(seq_len(ncol(.)))] } %>% graph_from_data_frame
  ## graph_from_data_frame does not care about column names,
  ## only about positions
  # %>%
  #graph_from_data_frame

## additional data:

graph_idxs <- match(var_names, V(hier_graph)$name)

hier_graph <- set_vertex_attr(hier_graph, "which_max_iso", graph_idxs,
                              dcor_subset %>% apply(1, which.max))
hier_graph <- set_vertex_attr(hier_graph, "dcor_max", graph_idxs,
                              dcor_subset %>% apply(1, max))
hier_graph <- set_vertex_attr(hier_graph, "dcor1", graph_idxs,
                              dcor_subset[, 1])
hier_graph <- set_vertex_attr(hier_graph, "dcor2", graph_idxs,
                              dcor_subset[, 2])
hier_graph <- set_vertex_attr(hier_graph, "dcor3", graph_idxs,
                              dcor_subset[, 3])
hier_graph <- set_vertex_attr(hier_graph, "dcor4", graph_idxs,
                              dcor_subset[, 4])
hier_graph <- set_vertex_attr(hier_graph, "dcor5", graph_idxs,
                              dcor_subset[, 5])
hier_graph <- set_vertex_attr(hier_graph, "dcor6", graph_idxs,
                              dcor_subset[, 6])
hier_graph <- set_vertex_attr(hier_graph, "dcor7", graph_idxs,
                              dcor_subset[, 7])
# 

hier_graph <-
  set_vertex_attr(hier_graph, "Indicator.Code", graph_idxs,
                  local_wdi_series$`Indicator Code`[
                               match(var_names, local_wdi_series$`Indicator Code`)])



# thisPeople<-fread("../data/indicator_compressed_names.csv")
# wdi_names_prepared<-wdi_series_used_important[,c("Indicator Code","Indicator Name")]
# 
# 
# merge(wdi_names_prepared,thisPeople,by="Indicator Code",all.x = T)->wdi_names_prepared2

# wdi_names_prepared2 %>% select(`Indicator Code`,`Indicator Name`,`Compressed Name`)%>%  filter(is.na(`Compressed Name`)) %>% fwrite("../output/preparethenameschatgpt.csv")
# 
# wdi_series_used[,c("Indicator Code","Compressed Name")] #%>% fwrite("../data/indicator_compressed_names.csv")
# # # chatgpt shortened the names
# wdi_series_used
# wdi_series_used2<-fread("../data/indicator_compressed_names.csv")
# shortnames<-fread("../data/05_abbrevationmaker.csv",sep=";")
# rbind(wdi_series_used2[!(wdi_series_used2$`Indicator Code`%in% shortnames$`Indicator Code`)],shortnames) %>%  fwrite("../output/05_indicator_compressed_names.csv")

wdi_series_used<-fread("../data/indicator_compressed_names.csv")

wdi_var_hierarchy_graph <-hier_graph
axis_cons <- lapply(1:7, inner_connections_by_axis,
                       dcor_var_iso_cons       = dcor_var_iso_cons,
                       wdi_var_hierarchy_graph = wdi_var_hierarchy_graph,
                       var_names               = var_names)




expand_multiplier = 1/2
text_size         = 3
label_modifier    = manual_label_modifier
#label_modifier    = short_name_label_modifier
## .overwrite = TRUE, # nolint: commented_code_linter.
.width  = 50
.height = 50


wdi_var_hierarchy_graph %>%
ggraph(layout = "dendrogram", circular = TRUE) +
geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor1),
                    y = xy_expand(y, dcor1)),
                colour = col_pal[1]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor1),
                 y = xy_expand(y, dcor1)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[1]) +
geom_node_point(aes(filter = leaf,
                   x = xy_expand(x, dcor2),
                    y = xy_expand(y, dcor2)),
                colour = col_pal[2]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor2),
                 y = xy_expand(y, dcor2)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[2]) +
geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor3),
                    y = xy_expand(y, dcor3)),
                colour = col_pal[3]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor3),
                 y = xy_expand(y, dcor3)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[3]) +
geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor4),
                    y = xy_expand(y, dcor4)),
                colour = col_pal[4]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor4),
                 y = xy_expand(y, dcor4)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[4]) +
geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor5),
                    y = xy_expand(y, dcor5)),
                colour = col_pal[5]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor5),
                 y = xy_expand(y, dcor5)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[5]) +
  geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor5),
                    y = xy_expand(y, dcor5)),
                colour = col_pal[5]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor5),
                 y = xy_expand(y, dcor5)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[5]) +
  geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor5),
                    y = xy_expand(y, dcor5)),
                colour = col_pal[6]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor5),
                 y = xy_expand(y, dcor5)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[6]) +
      geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor5),
                    y = xy_expand(y, dcor5)),
                colour = col_pal[7]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor5),
                 y = xy_expand(y, dcor5)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[7]) +
geom_node_point(aes(filter = leaf, color = factor(which_max_iso)),
                size = 0)+
geom_node_text(aes(filter = leaf,
                   label  = label_modifier(as.character(Indicator.Code)),
                   color  = factor(paste(component_common, which_max_iso)),
                   x      = x * (1 + 1 * expand_multiplier),
                   y      = y * (1 + 1 * expand_multiplier),
                   hjust  = if_else(x > 0, 0, 1),
                   angle  = atan(y / x) / 2 / pi * 360),
               show.legend = FALSE,
               size        = text_size) +
geom_conn_bundle(data = get_con(axis_cons[[1]][1, ], axis_cons[[1]][2, ]),
                 edge_colour = col_pal[1], edge_alpha = 0.3) +
geom_conn_bundle(data = get_con(axis_cons[[2]][1, ], axis_cons[[2]][2, ]),
                 edge_colour = col_pal[2], edge_alpha = 0.3) +
geom_conn_bundle(data = get_con(axis_cons[[3]][1, ], axis_cons[[3]][2, ]),
                 edge_colour = col_pal[3], edge_alpha = 0.3) +
(
  if (ncol(axis_cons[[4]]) > 0)
    geom_conn_bundle(data = get_con(axis_cons[[4]][1, ],
                                    axis_cons[[4]][2, ]),
                     edge_colour = col_pal[4], edge_alpha = 0.3)
  else
    list()
) +
geom_conn_bundle(data = get_con(axis_cons[[5]][1, ], axis_cons[[5]][2, ]),
                 edge_colour = col_pal[5], edge_alpha = 0.3) +
geom_edge_diagonal() +
# geom_node_label(aes(filter = !leaf,
#                     fontface = if_else(name %in% level1, "bold", "plain"),
#                     label = name, angle = atan(y / x) / 2 / pi * 360),
#                 size = 2) +
scale_color_manual(limits = component_names,
                   labels = component_names,
                   values = col_pal) +
guides(colour = guide_legend(title = component_common,
                             override.aes = list(size = 5))) +
theme_void() +
theme(legend.title = element_blank(),
      legend.position = c(0.9, 0.1),
      legend.justification = c(0.9, 0.1)) +
expand_limits(x = c(-2.7, 2.7), y = c(-2.7, 2.7))

ggsave("../figures/07_230808_circle_testshaplez.pdf",device = "pdf",height = 12,width = 12)

```


## PLOT 3 PCA

### input
#### data

```{r}
DATA_DIR<-"../data/"
RELATIVE_INDICATORS_FILE_NAME <- "../data/01_data/relative_indicators_2023_06.R"
importance_matrix <-fread("../data/07_importance_PCA.csv")

importance_matrix 
```


#### setup top 5
```{r}
cor_var_iso_cons <- fread("../data/08_wdi_pca_correlation.csv")
round(importance_matrix$Importance ,digits = 2)*100
importance_matrix

cor_var_PCA_cons<-cor_var_iso_cons[,colnames(cor_var_iso_cons)%in% (paste0(importance_matrix$Feature[1:7],".energy")),with=F] %>%  as.matrix()
rownames(cor_var_PCA_cons)<- cor_var_iso_cons$`Series Code`

```


```{r}
# nolint start
# used_var_names_max_cor_pca <- c(
#   cor_var_PCA_cons[,1 ] %>% order(decreasing = TRUE) %>% { .[1:24] } %>% { rownames(cor_var_PCA_cons)[.] },
#   cor_var_PCA_cons[,2] %>% order(decreasing = TRUE) %>% { .[1:18] } %>% { rownames(cor_var_PCA_cons)[.] },
#   cor_var_PCA_cons[,3] %>% order(decreasing = TRUE) %>% { .[1:14] } %>% { rownames(cor_var_PCA_cons)[.] },
#   cor_var_PCA_cons[, 4] %>% order(decreasing = TRUE) %>% { .[1:14] } %>% { rownames(cor_var_PCA_cons)[.] },
#   cor_var_PCA_cons[, 5] %>% order(decreasing = TRUE) %>% { .[1:10]  } %>% { rownames(cor_var_PCA_cons)[.] }
# ) %>% unique %>% sort




dim1<-  cor_var_PCA_cons[,1 ] %>% order(decreasing = TRUE) %>% { .[1:24] } %>% { rownames(cor_var_PCA_cons)[.] }
dim2<-  cor_var_PCA_cons[,2] %>% order(decreasing = TRUE) %>% { .[1:18] } %>% { rownames(cor_var_PCA_cons)[.] }
dim3<-  cor_var_PCA_cons[,3] %>% order(decreasing = TRUE) %>% { .[1:14] } %>% { rownames(cor_var_PCA_cons)[.] }
dim4<-  cor_var_PCA_cons[, 4] %>% order(decreasing = TRUE) %>% { .[1:14] } %>% { rownames(cor_var_PCA_cons)[.] }
dim5<-  cor_var_PCA_cons[, 5] %>% order(decreasing = TRUE) %>% { .[1:10]  } %>% { rownames(cor_var_PCA_cons)[.] }
dim6<-  cor_var_PCA_cons[, 6] %>% order(decreasing = TRUE) %>% { .[1:07] } %>% { names(cor_var_PCA_cons[., 6]) }
dim7<-  cor_var_PCA_cons[, 7] %>% order(decreasing = TRUE) %>% { .[1:06]  } %>% {names(cor_var_PCA_cons[., 7]) }
used_var_names_max_dcor <- c(dim1,dim2,dim3,dim4,dim5,dim6,dim7) %>% unique %>% sort

```


```{r}
subsetwdi<-wdi_explained[wdi_explained$`Series Code`%in% used_var_names_max_dcor,]
var_names         = used_var_names_max_dcor
dcor_var_iso_cons = cor_var_PCA_cons
load("../data/wdi_series")
local_wdi_series  = wdi_series[,]
expand_multiplier = 1 / 10
text_size         = 5
label_modifier    = manual_label_modifier
col_pal <- c("#1F968B", "#7570b3", "#3CBB75", "#e7298a", "yellow3", "orange","slategrey")


wdi_explained[wdi_explained$`Series Code`%in% used_var_names_max_dcor,] %>%  select("Topic","Indicator Name","Short definition" ) %>% fwrite("../output/listofimportantindicatores.csv")

```

#### modifiy the compressed names
```{r}
compressedNames <- fread("../data/indicator_compressed_names.csv")

missingIndicators <- wdi_explained %>%
  filter(`Series Code` %in% used_var_names_max_dcor) %>%
  select(`Series Code`,`Indicator Name`) %>%
  rename(`Indicator Code` = `Series Code`) %>%
  merge(compressedNames[,c("Indicator Code","Compressed Name")], by = "Indicator Code", all.x = TRUE) %>%
  filter(is.na(`Compressed Name`))

missingIndicators

# 
#missingIndicators %>% fwrite("../data/05_abbrevationmaker.csv",sep=";")

#rbind(compressedNames,addthis) %>% fwrite("../data/indicator_compressed_names.csv")
```



### plot function
```{r}

#label_modifier    = short_name_label_modifier,
## .overwrite = TRUE, # nolint: commented_code_linter.
.width  = 15
.height = 15

component_common <- "Dimension"
component_names <- paste(component_common, 1:7)
xy_expand <- function(xy, expand) xy * (1 + expand * expand_multiplier)
# Does not work
local_wdi_series<-local_wdi_series[grepl(":",local_wdi_series$Topic),]

local_wdi_series <- wdi_series_split_topic(local_wdi_series)

wdi_series_used_important <- local_wdi_series[`Indicator Code` %in% var_names]
var_names<-var_names[var_names%in% wdi_series_used_important$`Indicator Code`]
require(igraph)
stopifnot(length(var_names) == nrow(wdi_series_used_important))
dcor_var_iso_cons_subset<-dcor_var_iso_cons[rownames(dcor_var_iso_cons)%in% var_names,]
dcor_var_iso_cons<-dcor_var_iso_cons[rownames(dcor_var_iso_cons)%in% var_names,]



local_wdi_series<-local_wdi_series[local_wdi_series$`Indicator Code`%in% var_names,]
require(tidygraph)
require(ggraph)

dcor_subset<-dcor_var_iso_cons_subset
## local_wdi_series$`Indicator Code`
edges_level_01 <- local_wdi_series %>% select(`Indicator Code`, Topic1) %>%
  unique %>% rename( from = Topic1, to = `Indicator Code`)
edges_level_12 <- local_wdi_series %>% select(Topic1, Topic2)           %>%
  unique %>% rename( from = Topic2, to = Topic1          )
edges_level_23 <- local_wdi_series %>% select(Topic2, Topic3)           %>%
  unique %>% rename( from = Topic3, to = Topic2          )
edges_level_34 <- local_wdi_series %>% select(Topic3, Topic4)           %>%
  unique %>% rename( from = Topic4, to = Topic3          )
edges_level_45 <- local_wdi_series %>% select(Topic4, Topic5)           %>%
  unique %>% rename( from = Topic5, to = Topic4          )
## create the actual igraph object

rbind(edges_level_01, edges_level_12, edges_level_23, edges_level_34,
      edges_level_45) %>%
unique %>%
filter(to != from) ->listi
hier_graph <-listi %>% data.frame() %>%  { .[rev(seq_len(ncol(.)))] } %>% graph_from_data_frame
  ## graph_from_data_frame does not care about column names,
  ## only about positions
  # %>%
  #graph_from_data_frame

## additional data:

graph_idxs <- match(var_names, V(hier_graph)$name)
dcor_subset %>% data.frame()%>% rownames_to_column(var="indicator") %>% mutate(color_vector=case_when(indicator%in% dim7~7,indicator%in% dim6~6,indicator%in% dim5~5,indicator%in% dim4~4,indicator%in% dim3~3,indicator%in% dim2~2,indicator%in% dim1~1,TRUE~8))-> value_color

hier_graph <- set_vertex_attr(hier_graph, "which_max_iso", graph_idxs,value_color$color_vector)
                              #dcor_subset %>% apply(1, which.max))
hier_graph <- set_vertex_attr(hier_graph, "dcor_max", graph_idxs,
                              dcor_subset %>% apply(1, max))
hier_graph <- set_vertex_attr(hier_graph, "dcor1", graph_idxs,
                              dcor_subset[, 1])
hier_graph <- set_vertex_attr(hier_graph, "dcor2", graph_idxs,
                              dcor_subset[, 2])
hier_graph <- set_vertex_attr(hier_graph, "dcor3", graph_idxs,
                              dcor_subset[, 3])
hier_graph <- set_vertex_attr(hier_graph, "dcor4", graph_idxs,
                              dcor_subset[, 4])
hier_graph <- set_vertex_attr(hier_graph, "dcor5", graph_idxs,
                              dcor_subset[, 5])
hier_graph <- set_vertex_attr(hier_graph, "dcor6", graph_idxs,
                              dcor_subset[, 6])
hier_graph <- set_vertex_attr(hier_graph, "dcor7", graph_idxs,
                              dcor_subset[, 7])
# 

hier_graph <-
  set_vertex_attr(hier_graph, "Indicator.Code", graph_idxs,
                  local_wdi_series$`Indicator Code`[
                               match(var_names, local_wdi_series$`Indicator Code`)])



# thisPeople<-fread("../data/indicator_compressed_names.csv")
# wdi_names_prepared<-wdi_series_used_important[,c("Indicator Code","Indicator Name")]
# 
# 
# merge(wdi_names_prepared,thisPeople,by="Indicator Code",all.x = T)->wdi_names_prepared2

# wdi_names_prepared2 %>% select(`Indicator Code`,`Indicator Name`,`Compressed Name`)%>%  filter(is.na(`Compressed Name`)) %>% fwrite("../output/preparethenameschatgpt.csv")
# 
# wdi_series_used[,c("Indicator Code","Compressed Name")] #%>% fwrite("../data/indicator_compressed_names.csv")
# # # chatgpt shortened the names
# wdi_series_used
# wdi_series_used2<-fread("../data/indicator_compressed_names.csv")
# shortnames<-fread("../data/05_abbrevationmaker.csv",sep=";")
# rbind(wdi_series_used2[!(wdi_series_used2$`Indicator Code`%in% shortnames$`Indicator Code`)],shortnames) %>%  fwrite("../output/05_indicator_compressed_names.csv")

wdi_series_used<-fread("../data/indicator_compressed_names.csv")

wdi_var_hierarchy_graph <-hier_graph
axis_cons <- lapply(1:7, inner_connections_by_axis,
                       dcor_var_iso_cons       = dcor_var_iso_cons,
                       wdi_var_hierarchy_graph = wdi_var_hierarchy_graph,
                       var_names               = var_names)




expand_multiplier = 1/2
text_size         = 3
label_modifier    = manual_label_modifier
#label_modifier    = short_name_label_modifier
## .overwrite = TRUE, # nolint: commented_code_linter.
.width  = 50
.height = 50


wdi_var_hierarchy_graph %>%
ggraph(layout = "dendrogram", circular = TRUE) +
geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor1),
                    y = xy_expand(y, dcor1)),
                colour = col_pal[1]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor1),
                 y = xy_expand(y, dcor1)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[1]) +
geom_node_point(aes(filter = leaf,
                   x = xy_expand(x, dcor2),
                    y = xy_expand(y, dcor2)),
                colour = col_pal[2]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor2),
                 y = xy_expand(y, dcor2)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[2]) +
geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor3),
                    y = xy_expand(y, dcor3)),
                colour = col_pal[3]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor3),
                 y = xy_expand(y, dcor3)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[3]) +
geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor4),
                    y = xy_expand(y, dcor4)),
                colour = col_pal[4]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor4),
                 y = xy_expand(y, dcor4)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[4]) +
geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor5),
                    y = xy_expand(y, dcor5)),
                colour = col_pal[5]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor5),
                 y = xy_expand(y, dcor5)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[5]) +
  geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor5),
                    y = xy_expand(y, dcor5)),
                colour = col_pal[5]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor5),
                 y = xy_expand(y, dcor5)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[5]) +
  geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor5),
                    y = xy_expand(y, dcor5)),
                colour = col_pal[6]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor5),
                 y = xy_expand(y, dcor5)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[6]) +
      geom_node_point(aes(filter = leaf,
                    x = xy_expand(x, dcor5),
                    y = xy_expand(y, dcor5)),
                colour = col_pal[7]) +
geom_polygon(aes(sort = name,
                 filter = name %in% var_names,
                 x = xy_expand(x, dcor5),
                 y = xy_expand(y, dcor5)),
             stat = stat_filter_sort,
             group = 1,
             fill = NA,
             colour = col_pal[7]) +
geom_node_point(aes(filter = leaf, color = factor(which_max_iso)),
                size = 0)+
geom_node_text(aes(filter = leaf,
                   label  = label_modifier(as.character(Indicator.Code)),
                   color  = factor(paste(component_common, which_max_iso)),
                   x      = x * (1 + 1 * expand_multiplier),
                   y      = y * (1 + 1 * expand_multiplier),
                   hjust  = if_else(x > 0, 0, 1),
                   angle  = atan(y / x) / 2 / pi * 360),
               show.legend = FALSE,
               size        = text_size) +
geom_conn_bundle(data = get_con(axis_cons[[1]][1, ], axis_cons[[1]][2, ]),
                 edge_colour = col_pal[1], edge_alpha = 0.3) +
geom_conn_bundle(data = get_con(axis_cons[[2]][1, ], axis_cons[[2]][2, ]),
                 edge_colour = col_pal[2], edge_alpha = 0.3) +
geom_conn_bundle(data = get_con(axis_cons[[3]][1, ], axis_cons[[3]][2, ]),
                 edge_colour = col_pal[3], edge_alpha = 0.3) +
(
  if (ncol(axis_cons[[4]]) > 0)
    geom_conn_bundle(data = get_con(axis_cons[[4]][1, ],
                                    axis_cons[[4]][2, ]),
                     edge_colour = col_pal[4], edge_alpha = 0.3)
  else
    list()
) +
geom_conn_bundle(data = get_con(axis_cons[[5]][1, ], axis_cons[[5]][2, ]),
                 edge_colour = col_pal[5], edge_alpha = 0.3) +
geom_edge_diagonal() +
# geom_node_label(aes(filter = !leaf,
#                     fontface = if_else(name %in% level1, "bold", "plain"),
#                     label = name, angle = atan(y / x) / 2 / pi * 360),
#                 size = 2) +
scale_color_manual(limits = component_names,
                   labels = component_names,
                   values = col_pal) +
guides(colour = guide_legend(title = component_common,
                             override.aes = list(size = 5))) +
theme_void() +
theme(legend.title = element_blank(),
      legend.position = c(0.9, 0.1),
      legend.justification = c(0.9, 0.1)) +
expand_limits(x = c(-2.7, 2.7), y = c(-2.7, 2.7)) + scale_fill_discrete(labels=paste(importance_matrix$Feature[1:7]))
#ggsave("../figures/07_2308622_pca.pdf",device = "pdf",height = 12,width = 12)

```


# Result 6 THE round

Conrol THE outliers
```{r}
#subsetwdi
num_levels <- 4
# Use cut() to create levels of quantiles
dimred2021_knn250_2$pscore_level<-as.factor(ntile(dimred2021_knn250_2$pscore_mean, num_levels))
dimred2021_knn250_2 %>% group_by(pscore_level) %>% summarize(Min=min(pscore_mean),Max=max(pscore_mean)) %>% mutate(p_score_level_label=paste0(Min," to ",Max)) %>% 
  left_join(dimred2021_knn250_2, . ,by="pscore_level")-> dimred2021_knn250_3
dimred2021_knn250_3 %>%filter(year%in% 2020) %>%   select(`Country Code`,`Short Name`,TX.VAL.MRCH.R3.ZS,Region.x) %>%  arrange(desc(TX.VAL.MRCH.R3.ZS))
dimred2021_knn250_3 %>% filter(year%in% 2020) %>%  select(`Country Code`,`Short Name`,SL.TLF.CACT.FE.ZS,Region.x) %>%  arrange(desc(SL.TLF.CACT.FE.ZS))

```


```{r}
subsetwdi<-wdi_explained[wdi_explained$`Series Code`%in% colnames(dimred2021_knn250_5),]
dimred2021_knn250_5 %>% 
  # mutate(pscore_cat2 = fct_relevel(pscore_cat, 
  #           "", "<0", "0 to <5","5 to <10","10 to <20","20 to <30","30 to <40","40 to <50", ">=50" )) %>% 
  mutate(p_score_level_label = fct_relevel(p_score_level_label,
            "-6.4 to 5.5" , "5.5 to 8.6" , "8.6 to 18.1" ,"18.1 to 97.1"  )) %>%
  mutate(p_score_level_label2=case_when(p_score_level_label%in% "18.1 to 97.1" & RegionSoviet %in% "Post-Soviet"~"High Soviets",
                                            p_score_level_label%in% "18.1 to 97.1" & RegionSoviet %in% "Latin America & Caribbean"~"High Latin",
                                               p_score_level_label%in% "18.1 to 97.1" ~"High other",
                                              TRUE~p_score_level_label ))    %>%                                                                                     
  # mutate(p_score_level_label = fct_relevel(p_score_level_label, 
  #           "-6.4 to 4.4","4.4 to 7" ,  "7 to 11.6" ,"11.6 to 20.1" ,"20.1 to 97.1" )) %>% 

  filter(!is.na(pscore_level))->dimred2021_knn250_3 
dimred2021_knn250_3$p_score_level_label2 %>%  table
levels(as.factor(dimred2021_knn250_3$p_score_level_label))
Indicator<-"EN.ATM.CO2E.KD.GD"
year_start<-"2015"
year_end<-"2021"

for(i in 1:nrow(subsetwdi)){
  Indicator<-subsetwdi[i,"Series Code"] %>%  unlist() %>%  as.character()
  
  a<-get_wdi_trajectory(data=dimred2021_knn250_3,Indicator,groupedby ="p_score_level_label2" ,CI = T)
  a %>% ggsave(file = paste0("../test/trajectories/split_regions_all/",str_replace_all(Indicator,"\\.","_"),".png"),device = "png")
}
dimred2021_knn250_5 %>%  filter(p_score_level_label2%in% "High other") %>%  select(`Country Code`,RegionSoviet) %>%  unique ->sad
sad$RegionSoviet %>%  table
dimred2021_knn250_5 %>%  filter(p_score_level_label2%in% "High other") %>% select(`Country Code`,RegionSoviet,`Country Name`) %>%  unique-> asd
```


```{r}
a<-get_wdi_trajectory(data=dimred2021_knn250_3,"iso1","p_score_level_label2",CI = T)
b<-get_wdi_trajectory(data=dimred2021_knn250_3,"iso2","p_score_level_label2",CI = T)
c<-get_wdi_trajectory(data=dimred2021_knn250_3,"iso3","p_score_level_label2",CI = T)
d<-get_wdi_trajectory(data=dimred2021_knn250_3,"iso4","p_score_level_label2",CI = T)

ggarrange(a+theme(legend.position = "asd")+xlab("")+ggtitle("iso1"),b+theme(legend.position = "asd")+xlab("")+ggtitle("iso2"),c+theme(legend.position = "asd")+xlab("")+ggtitle("iso3"),d+theme(legend.position = "asd")+xlab("")+ggtitle("iso4")) %>%  ggsave(file="../figures/07_230810_isoExcess.png",device = "png")
```
#### High former Soviet statets

add indicators that are not in the dimred
```{r}
iso2allcors<-fread("../results/corrTable/05_CorrTable_iso2_quantiles_shapleys_2019to2021_all.csv")
iso2allcors$iso2_all %>%  summary()
wdi_data<-fread("../data/05_WDI_data.csv")


add_col<-c(colnames(wdi_data)[!(colnames(wdi_data)%in% colnames(dimred2021_knn250_3))])

wdi_data %>% mutate(identifier=paste0(`Country Code`,year)) %>%select(any_of(c("Country Code","year","identifier",add_col)))-> wdi_data2
dimred2021_knn250_3 %>% mutate(identifier=paste0(`Country Code`,year)) -> dimred2021_knn250_4

left_join(dimred2021_knn250_4,wdi_data2 %>% select(any_of(c("identifier",add_col))),"identifier") %>%  filter() -> dimred2021_knn250_5

iso2allcors %>% filter(N_75to100_75to100>mean(N_75to100_75to100)) %>%  arrange(desc(iso2_75to100_75to100)) %>% slice(1:100)-> newsubset
newsubset->newsubset
for(i in 1:nrow(newsubset)){
  Indicator<-newsubset[i,"Series Code"] %>%  unlist() %>%  as.character()
  a<-get_wdi_trajectory(data=dimred2021_knn250_5,Indicator,groupedby ="p_score_level_label2" ,CI = T)
  
  a %>% ggsave(file = paste0("../test/trajectories/ISO2_forsoviets/",str_replace_all(Indicator,"\\.","_"),".png"),device = "png")
}
  a<-get_wdi_trajectory(data=dimred2021_knn250_5,"SP.POP.DPND.OL",groupedby ="p_score_level_label2" ,CI = T)
a
dimred2021_knn250_5 %>% filter(year%in% 2020)%>%  arrange(desc(SP.POP.DPND.OL)) %>% select(RegionSoviet,iso1,iso2,iso3,iso4)
```















# SUPPLEMENT

# 3 Figure nonlinear isomap types

```{r}
wdi_data_cons_df<-wdi_data_cons_df

featureList<-c("NY.GDP.PCAP.CD","Country Code","P_mort_2020","P_mort_2021","iso1","iso2","iso3","iso4","iso5","iso6","iso7","iso8","iso9","iso10","year","pscore_mean","RegionSoviet","Income Group")
```


## Plot gdp iso 1
```{r}
subset2020<-dimred2021_knn250_2 %>% filter(year%in% 2020) %>% select(any_of(featureList)) %>% rename(ExcessMort=P_mort_2020) %>%  mutate(P_mort_2020=NULL)
subset2021<-dimred2021_knn250_2 %>% filter(year%in% 2021) %>% select(any_of(featureList)) %>% rename(ExcessMort=P_mort_2021) %>% mutate(P_mort_2021=NULL)

bothyears<-rbind(subset2020,subset2021,use.names=FALSE)


bothyears %>% filter(year%in% 2021) %>% mutate(Income_group=`Income Group`) %>% filter(!is.na(Income_group))%>% 
  mutate(Income_group = factor(Income_group, levels=c("High income",
                                      "Upper middle income",
                                      "Lower middle income",
                                      "Low income"
                                      )),`GDP per Capita (log)`=log(NY.GDP.PCAP.CD),
         `Isomap dimension 1`=iso1
         ) %>% 
  
   pivot_longer(names_to = "var",
               c("Isomap dimension 1","GDP per Capita (log)"),
               values_drop_na = TRUE) %>% 
  ggplot(aes(x=(value),y=ExcessMort,color=Income_group))+geom_point()+ theme_bw()+xlab("")+ylab("Excess Mortality") +scale_color_viridis(discrete = T)+facet_wrap(~var,scales = "free_x")+theme(legend.position = "bottom") 
#ggsave("../figures/06_figures_GDP_Iso1.png",device = "png", width = 7, height = 5, dpi = 300, units = "in")
```
## Plot iso 2

### check all the plots
```{r}

wdi_dimred_iso2 %>% filter(!is.na(iso2_all))%>%  mutate(maxdisthighlow=abs(iso2_0to25-iso2_50to75),
                                                        adjustvalue=iso2_75to100/maxdisthighlow) %>% arrange(desc(maxdisthighlow))-> subsetiso2

# for(i in 1:676){
# nameVar<-subsetiso2[i,"Series Code"] %>%  unlist
# applySD<-!is.na(dimred2021_knn250_2[,..nameVar]) %>%  unlist
# 
# wool<-dimred2021_knn250_2[applySD[,1]] %>% filter(year%in% 2010:2021) %>% filter(!is.na(iso2)) %>% 
# ggplot(aes_string(x="iso2",y=paste(nameVar)))+geom_point()
# wool %>% ggsave(file = paste0("../test/run",i,".png",sep=""),device = "png")
# }

```
#### candidates
```{r}
iso2_candidates<-c(5,15,21,26,27,32,49,57,66,68,70,71,80,97,102,119,123,180,207,377,383,407,439,583)
nameVar_iso2_candidates<-subsetiso2[iso2_candidates,"Series Code"] 

for(i in iso2_candidates){
nameVar<-subsetiso2[i,"Series Code"] %>%  unlist
applySD<-!is.na(dimred2021_knn250_2[,..nameVar]) %>%  unlist

wool<-dimred2021_knn250_2[applySD[,1]] %>% filter(year%in% 2010:2021) %>% filter(!is.na(iso2))


wool %>%  mutate(categorie=case_when(scale(iso2)>1.5~"Iso2 peak  high",
                                                 scale(iso2)<  -0.5~"Iso2 peak  Low",
                                                 (scale(iso2)> (-0.5) &scale(iso2)<1.5)~"normal",
                                                  TRUE~"a"
                                                  ))%>% 
  filter(!is.na(get(nameVar)),!is.na(categorie)) -> yo


yo$categorie<-factor(as.factor(yo$categorie),levels=c("Iso2 peak  Low","normal","Iso2 peak  high","a"))
yo2<-yo[,colnames(yo)[colnames(yo)%in% c(paste(nameVar),"categorie")],with=F]
colnames(yo2)[1]<-"value"
series_label<-wdi_explained[wdi_explained$`Series Code` %in% nameVar,"Indicator Name"]

wool2<-ggstatsplot::ggbetweenstats(
  plot.type = "boxviolin",
                            p.adjust.method = "BH",
        point.args = list(alpha = 0),

      data = yo2,
      x = categorie,
      y = value ) + 
      # Add labels and title
      labs(x="",
           y = nameVar,
           title = series_label
      )+theme(
        # This is the new default font in the plot
        text = element_text(family = "Roboto", size = 10, color = "black"),
        plot.title = element_text(
          family = "Lobster Two", 
          size = 14,
          face = "bold",
          color = "#2a475e"
        ),
        # # Statistical annotations below the main title
        plot.subtitle = element_text(
          family = "Roboto",
          face = "bold",
          color="#1b2838"
        ),
        plot.title.position = "plot", # slightly different from default
        axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 12)
      )

wool2 %>% ggsave(file = paste0("../figures/06_iso2plots/",i,".png",sep=""),device = "png")
}

```

### plots
```{r}
featureList<-c("SH.STA.OWGH.ME.ZS","VC.IHR.PSRC.MA.P5","TM.VAL.MRCH.R3.ZS","TX.VAL.MRCH.R2.ZS","SP.DYN.CDRT.IN","SP.POP.2024.FE.5Y","Country Code","P_mort_2020","P_mort_2021","iso1","iso2","iso3","iso4","iso5","iso6","iso7","iso8","iso9","iso10","year","pscore_mean","RegionSoviet","Income Group")


#
# 
# ggsave("../figures/06_figuressa_GDP_Iso1.png",device = "png", width = 7, height = 5, dpi = 300, units = "in")
```
```{r}
dimred2021_knn250_2 %>% filter(year%in% 2019:2021) %>% mutate(Income_group=`Income Group`) %>% filter(!is.na(Income_group)) %>% 
  ggplot(aes(x=(iso2),y=(SP.DYN.CDRT.IN)))+geom_point()+geom_smooth(method = "lm")+ theme_bw()+xlab("Dimension 2")+ylab("Death rate, crude (per 1,000 people)") +scale_color_viridis(discrete = T) 


```
```{r}
dimred2021_knn250_2 %>% filter(year%in% 2019:2021) %>% mutate(Income_group=`Income Group`) %>% filter(!is.na(Income_group)) %>% 
  ggplot(aes(x=(iso2),y=TM.VAL.MRCH.R3.ZS))+geom_point()+geom_smooth(method = "lm")+ theme_bw()+xlab("Dimension 2")+ylab("Merchandise imports from low- and middle-income economies in Latin America & the Caribbean (% of total merchandise imports)") +scale_color_viridis(discrete = T) 
dimred2021_knn250_2 %>% filter(year%in% 2019:2021) %>% mutate(Income_group=`Income Group`) %>% filter(!is.na(Income_group)) %>% 
  ggplot(aes(x=(iso2),y=TX.VAL.MRCH.R2.ZS))+geom_point()+geom_smooth(method = "lm")+ theme_bw()+xlab("Dimension 2")+ylab("Merchandise exports to low- and middle-income economies in Europe & Central Asia (% of total merchandise exports)") +scale_color_viridis(discrete = T) 
```

## Plot iso 1

### check all the plots
```{r}
wdi_dimred_iso1<-fread("../results/corrTable/04_CorrTable_iso1_quantiles_all.csv")
wdi_dimred_iso1 %>% filter(!is.na(iso1_all))%>%  mutate(maxdisthighlow=abs(iso1_0to25-iso1_50to75),
                                                        adjustvalue=iso1_75to100/maxdisthighlow) %>% arrange(desc(maxdisthighlow))-> subsetiso1
```


```{r}
for(i in 1:100){
nameVar<-subsetiso1[i,"Series Code"] %>%  unlist
applySD<-!is.na(dimred2021_knn250_2[,..nameVar]) %>%  unlist

wool<-dimred2021_knn250_2[applySD[,1]] %>% filter(year%in% 2010:2021) %>% filter(!is.na(iso2)) %>%
ggplot(aes_string(x="iso1",y=paste(nameVar)))+geom_point()
wool %>% ggsave(file = paste0("../test/run",i,".png",sep=""),device = "png")
}

```
#### candidates
```{r}
iso1_candidates<-c(24,3,34,58)
nameVar_iso1_candidates<-subsetiso1[iso1_candidates,"Series Code"] 

indicator_a<-nameVar_iso1_candidates$`Series Code`[1]

a<-dimred2021_knn250_2 %>% filter(year%in% 2000:2021) %>% 
  ggplot(aes_string(x="iso1",y=indicator_a))+geom_point(color="firebrick4")+ggtitle(paste(wdi_explained[wdi_explained$`Series Code`%in% indicator_a,"Indicator Name"]))+theme_bw()

indicator_b<-nameVar_iso1_candidates$`Series Code`[2]

b<-dimred2021_knn250_2 %>% filter(year%in% 2000:2021) %>% 
  ggplot(aes_string(x="iso1",y=indicator_b))+geom_point(color="blue4")+ggtitle(paste(wdi_explained[wdi_explained$`Series Code`%in% indicator_b,"Indicator Name"]))+theme_bw()
ggarrange(a,b,ncol = 1) %>% ggsave(file="../figures/06_figures_iso1_nonlinear.png",device = "png", width = 7, height = 5, dpi = 300, units = "in")
```

## Plot iso 4

### check all the plots
```{r}
wdi_dimred_iso4<-fread("../results/corrTable/04_CorrTable_iso4_quantiles_all.csv")
wdi_dimred_iso4 %>% filter(!is.na(iso4_all_all))%>%  mutate(maxdisthighlow=abs(iso4_0to25-iso4_75to100),
                                                        adjustvalue=iso4_50to75/maxdisthighlow) %>% arrange(desc(maxdisthighlow))-> subsetiso4
```


```{r}
for(i in 1:100){
nameVar<-subsetiso4[i,"Series Code"] %>%  unlist
applySD<-!is.na(dimred2021_knn250_2[,..nameVar]) %>%  unlist

wool<-dimred2021_knn250_2[applySD[,1]] %>% filter(year%in% 2010:2021) %>% filter(!is.na(iso4)) %>%
ggplot(aes_string(x="iso4",y=paste(nameVar)))+geom_point()
wool %>% ggsave(file = paste0("../test/run",i,".png",sep=""),device = "png")
}

```
#### candidates
```{r}
iso1_candidates<-c(24,3,34,58)
nameVar_iso1_candidates<-subsetiso1[iso1_candidates,"Series Code"] 

indicator_a<-nameVar_iso1_candidates$`Series Code`[1]

dimred2021_knn250_2 %>% filter(year%in% 2000:2021) %>% 
  ggplot(aes_string(x="iso1",y=indicator_a))+geom_point()+ggtitle(paste(wdi_explained[wdi_explained$`Series Code`%in% indicator_a,"Indicator Name"]))
```


# 4 Shapley values

